% Project Ghost — arXiv submission (v0.2.3, 2026-06-10)
% Author: Javier Menéndez Mateos <jfhelvetius@gmail.com>
% Compiles with pdflatex + bibtex (or biber if you switch \bibliographystyle).
%
%   pdflatex main
%   bibtex main
%   pdflatex main
%   pdflatex main
%
% This source is also the basis for RV 2026 / FMAS 2026 workshop
% submissions; swap \documentclass to the venue's required template
% (e.g. lncs.cls for Springer LNCS workshops).

\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{microtype}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,shapes.geometric,backgrounds,fit}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[colorlinks=true,linkcolor=blue!60!black,citecolor=blue!50!black,urlcolor=blue!60!black]{hyperref}

% --- Listings (Python / shell / JSON) -----------------------------------
\definecolor{codebg}{rgb}{0.97,0.97,0.97}
\lstset{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{codebg},
  frame=single,
  rulecolor=\color{black!20},
  numbers=none,
  breaklines=true,
  showstringspaces=false,
  columns=fullflexible,
  xleftmargin=0.5em,
  xrightmargin=0.5em,
}

% --- Theorem environments -----------------------------------------------
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}

% --- Title block --------------------------------------------------------
\title{Epistemic Contracts for Autonomous Systems:\\
       A Verifiable Pattern for Safety Claims Under Uncertainty}
\author{Javier Menéndez Mateos\\
        \texttt{jfhelvetius@gmail.com}\\
        \small Independent}
\date{v0.2.3 --- \today}

\begin{document}
\maketitle

\begin{center}
\itshape\small
``An autonomous agent should have verifiable obligations about
how it relates to its own uncertainty.''\\[0.3em]
``A safety claim should be issued together with everything a third
party needs to reject it.''\\[0.3em]
\upshape\textmd{--- the two sentences this paper exists to defend.}
\end{center}

\vspace{1em}

\begin{abstract}
\noindent\textbf{Background.} Existing runtime verifiers
predominantly monitor predicates over the external world
(velocity, distance, temperature). What they do not monitor
are predicates over the \emph{agent's own posture under
uncertainty} --- how it degrades confidence, recovers when
evidence is restored, remains bounded in latency, and
translates belief into action.

\noindent\textbf{Aims.} We propose \textbf{epistemic safety
contracts} as a property class distinct from STL-style
predicates over signals and POMDP-style belief monitoring,
formalise the class, and exhibit a working implementation,
mechanical proofs, and a real-telemetry experiment.

\noindent\textbf{Method.} We describe \textbf{Project Ghost},
an open-source platform that (i) defines \textbf{seven}
epistemic contracts for a reference autonomy supervisor
(BAUD-v1, ERUR-v1, ERUR-v2, MD-v1, RLB-v1, FPB-v1, FPB-v2),
formalised as a Python \texttt{Protocol} and registry
(ADR-0045); (ii) verifies each via a pure function over a
content-addressed MCAP log; (iii) mechanically checks the
underlying invariants via TLA+/TLC at bounded $W$, a
parametric TLC sweep over $W \in \{4, 8, 16\}$, and Lean 4
proofs (partition theorem fully discharged, no \texttt{sorry};
RLB-v1 unbounded mechanically reduced to a single
load-bearing lemma); and (iv) closes the previously-open
``Python $\leftrightarrow$ TLA+ bridge by inspection'' caveat
with a Hypothesis-checked conformance test (ADR-0043). Each
contract ships as an \textbf{executable safety citation}:
\texttt{pip install project-ghost} followed by
\texttt{ghost verify-properties --mcap <log>} lets a third
party reproduce the verdict --- or contradict it.

\noindent\textbf{Results.} Empirical evaluation on a violation
matrix of six injected bug categories, three structurally
distinct calibration policies, three shape-realistic drift
profiles, a head-to-head benchmark against RTAMT, and a
\textbf{three-ULog discrimination experiment} on real PX4 SITL
flight telemetry establishes that the verifier is
policy-agnostic, deterministic across Linux and Windows CI
runners, and informative on real-world telemetry. With
independent simulator GT auto-detected (ADR-0037, v0.2.5), the
$3 \times 6 = 18$ discrimination matrix is $18/18$ green;
$15/18$ cells isolate the violation to the expected property
(the single non-isolated row is a documented true co-violation
between BAUD-v1 and MD-v1).

\noindent\textbf{Conclusions.} Epistemic safety contracts
constitute a tractable property class for autonomous agents.
The framework supports incremental contract addition (one
\texttt{register\_contract(...)} call), keeps verifiers
backward-compatible, and produces mechanical evidence that
withstands third-party audit. The full artifact is re-runnable
from a fresh clone (see [AUDIT.md, REPRODUCE.md, INSTALL.md]).
\end{abstract}

\medskip
\noindent\textbf{Keywords:} epistemic safety contracts, runtime
verification, uncertainty in autonomy, executable safety citations,
content-addressed telemetry, TLA+/TLC, MCAP.

\medskip
\noindent\textbf{Artifact availability:}
\begin{itemize}[noitemsep,topsep=0pt]
  \item Source: \url{https://github.com/JFHelvetius/ghost}
  \item Distribution: \url{https://pypi.org/project/project-ghost/}
  \item Documentation: \url{https://JFHelvetius.github.io/ghost/}
  \item License: Apache-2.0
\end{itemize}

% =======================================================================
\section{Introduction}
\label{sec:intro}

Most existing runtime verifiers ask predicates of the world:
velocity below a bound, distance above a margin, temperature
inside an envelope. We ask predicates of the agent's posture
toward its own uncertainty: contracts the agent must satisfy
about \emph{how} it degrades confidence, \emph{how} it recovers,
\emph{how} its uncertainty stays bounded, and \emph{how} belief
is translated into action. We call these \textbf{epistemic safety
contracts}.

Epistemic contracts are not contracts about what the agent
believes; they are contracts about what the agent must do
\emph{given} what it believes about itself.
``If you detect that your calibration history contains evidence of
drift, you may not emit a non-conservative action'' (BAUD) is a
different shape of property than ``velocity must remain below
5\,m/s'' (an STL-style predicate over a signal): the precondition
refers to the agent's self-assessment, not to the world.

A second gap follows: even if the right shape of property exists,
a third party who wants to verify a safety claim against a
recorded run typically cannot --- there is no shell command, no
content-addressed log, no pure-function verifier they can re-run
on their own machine. We close both gaps in one platform. Project
Ghost is sim-first, written in Python, and ships as a
\texttt{pip}-installable package with a CLI subcommand
(\texttt{ghost verify-properties}) that takes a captured MCAP log
and returns a byte-exact verdict on five epistemic contracts for
a reference autonomy supervisor. Each contract is stated as a
binding ADR, verified by a pure function over the log, exercised
by Hypothesis-based property tests, and self-enforced on every
push by CI. Two contracts (BAUD-v1 and ERUR-v1) are additionally
\textbf{mechanically verified} by TLA+/TLC, along with the
partition theorem that the two together cover the full
conditional behaviour space.

The packaging of an epistemic contract together with its recorded
run and its verifier into one citable, third-party-falsifiable
unit is what we call an \textbf{executable safety citation}. The
cited artefact \emph{is} the falsification mechanism.

\subsection{Contributions}
\label{sec:contributions}

\textbf{Epistemic safety contracts are verifiable obligations an
autonomous agent must satisfy about its own uncertainty.} A
contract is a triple (precondition over the agent's epistemic
state, postcondition over the agent's behaviour, pure-function
verifier over a recorded run) --- a different property class from
the world-predicates that dominate runtime verification today. We
package each contract together with the run and the verifier into
an \textbf{executable safety citation}.

We make \textbf{three contributions}:

\begin{itemize}
  \item \textbf{C1 --- Epistemic safety contracts as a
        verification target (conceptual).} A class of safety
        properties whose preconditions refer to the agent's
        belief about its own uncertainty
        (calibrated-self-assessment level, drift detection,
        fire-rate measurement) rather than to signals of the
        external world. Distinct from STL-style predicates over
        signals and from POMDP-style belief monitoring; the
        formal definition is in \S\ref{sec:epistemic-def}.
  \item \textbf{C2 --- Reference implementation: Ghost
        (artefact).} A closed-loop autonomy supervisor
        instantiating five epistemic contracts (BAUD-v1, ERUR-v1,
        MD-v1, RLB-v1, FPB-v1) packaged as executable safety
        citations --- binding ADRs, content-addressed MCAP
        telemetry, \texttt{ghost verify-properties --mcap},
        OIDC-signed PyPI wheels, and a policy-agnostic verifier
        across three calibration policies
        (\S\ref{sec:multi-policy}).
  \item \textbf{C3 --- Mechanical verification + empirical
        evaluation (validation).} Three TLA+ specifications
        (11 invariants in CI, including the partition theorem
        $\mathrm{BAUD} \oplus \mathrm{ERUR}$), a six-category
        violation matrix (\S\ref{sec:violation}), shape-realistic
        drift profiles, an RTAMT capability benchmark, and
        discrimination on real PX4 flight telemetry: two buggy
        components, swapped into the same physical flight, each
        flip BAUD-v1 from HOLDS to VIOLATED
        (\S\ref{sec:realflight-discrim}).
\end{itemize}

The recovery latency bound $L \le \mathrm{peak} + W - 1$ is
presented as a \textbf{supporting result}, not as a contribution.
Figure~\ref{fig:pattern} summarises the pattern.

\subsection{Epistemic safety contracts: formal definition}
\label{sec:epistemic-def}

An \textbf{epistemic safety contract} is a triple $(P, Q, V)$
where:

\begin{itemize}[noitemsep,topsep=2pt]
  \item $P$ is a predicate over the agent's \emph{epistemic state
        at cycle $t$} --- a function of the agent's self-assessment
        record, calibration history, and outcome stream available
        at $t$. $P$ does not refer directly to the world; it
        refers to what the agent believes about its own posture
        toward the world.
  \item $Q$ is a predicate over the agent's \emph{behaviour at
        cycle $t$} --- a function of the calibrated assessment,
        decision, and actuator command emitted at $t$.
  \item $V$ is a pure-function verifier such that, given a
        recorded run $r$ (a content-addressed sequence of
        per-cycle telemetry records), $V(r)$ returns HOLDS iff
        every cycle $t$ of $r$ that satisfies $P_t$ also satisfies
        $Q_t$, and VIOLATED otherwise with a witness cycle.
\end{itemize}

Three observations follow:

\begin{enumerate}[noitemsep,topsep=2pt]
  \item \textbf{An epistemic contract is not an STL predicate.}
        STL monitors evaluate predicates of the form
        \texttt{signal < threshold} over real-valued temporal
        signals. The atomic predicates of an epistemic contract
        are over the agent's \emph{internal record} of its own
        calibration and decisions, which is structured data, not
        a continuous signal. STL operators (always, eventually,
        until) can be lifted to act over the cycle index of an
        epistemic contract; this is how RLB-v1 (a bounded
        eventuality) is expressed.
  \item \textbf{An epistemic contract is not belief monitoring.}
        Belief monitoring tracks what an agent believes about
        hidden world state (e.g.\ POMDP belief updates,
        particle-filter posteriors). An epistemic contract
        verifies an obligation about the agent's \emph{relation
        to} its own belief --- that it must downgrade confidence
        in the right conditions, recover within bounded latency,
        and never claim more confidence than evidence supports.
        The contract sits one level above belief: it is a
        property of the agent's epistemic \emph{policy}, not of
        its beliefs themselves.
  \item \textbf{An epistemic contract becomes a
        third-party-falsifiable safety claim when packaged
        together with a content-addressed recorded run and an
        OIDC-signed verifier wheel.} We call this packaging an
        \textbf{executable safety citation}
        (Figure~\ref{fig:pattern}). The citation pattern is the
        mechanism by which an epistemic contract is shipped as
        something a reviewer can falsify in one shell command
        rather than as prose.
\end{enumerate}

The five contracts we ship (\S\ref{sec:properties}) instantiate
this definition on a representative autonomy supervisor. The
framework permits other contracts; FPB-v1 can be tightened, and
ports to other agent kinds (delivery, manipulation, multi-agent)
are scope for future work.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  font=\footnotesize,
  node distance=4mm and 6mm,
  box/.style={
    draw, rounded corners=2pt, align=center, inner sep=4pt,
    minimum width=29mm, minimum height=10mm
  },
  producer/.style={box, fill=blue!8, draw=blue!50!black},
  ci/.style={box, fill=green!10, draw=green!50!black},
  artifact/.style={box, fill=yellow!18, draw=orange!70!black},
  party/.style={box, fill=red!8, draw=red!60!black},
  flow/.style={-{Latex[length=2mm,width=1.6mm]}, thick},
  flowdash/.style={-{Latex[length=2mm,width=1.6mm]}, thick, dashed, gray!70!black},
  flowbig/.style={
    -{Latex[length=2.6mm,width=2mm]}, ultra thick, draw=red!50!black
  },
]

\node[producer] (adr) {Binding ADR\\\emph{property predicate}};
\node[producer, below=of adr] (code) {Pure-function verifier\\+ closed-loop pipeline};
\node[producer, below=of code] (spec) {Hypothesis tests\\+ TLA+ (11 invariants)};

\node[ci, right=14mm of code] (ci) {CI gate\\verify + TLC + cross-machine};
\node[ci, below=of ci] (tag) {Tagged release\\OIDC-signed PyPI wheel};

\node[artifact, right=14mm of ci] (mcap) {MCAP log\\SHA-256, byte-exact};
\node[artifact, below=of mcap] (cite) {\textbf{The citation}\\\texttt{pip install} + SHA-256};

\node[party, right=14mm of mcap] (cmd) {\texttt{ghost verify-}\\\texttt{properties --mcap}};
\node[party, below=of cmd] (out) {Exit 0/1 +\\ deterministic JSON};

% producer chain
\draw[flow] (adr) -- (code);
\draw[flow] (code) -- (spec);

% producer -> CI
\draw[flow] (spec.east) -- (ci.west);
\draw[flow] (ci) -- (tag);

% producer -> artifact (the run)
\draw[flowdash] (code.east) -- ++(3mm,0) |- (mcap.west);

% release + run -> citation
\draw[flow] (tag.east) -- ++(3mm,0) |- (cite.west);
\draw[flow] (mcap) -- (cite);

% citation -> third party
\draw[flowbig] (cite.east) -- ++(3mm,0) |- (cmd.west);
\draw[flow] (cmd) -- (out);

% Cluster labels above each subgraph
\node[font=\scriptsize\bfseries, above=2mm of adr] {Producer};
\node[font=\scriptsize\bfseries, above=2mm of ci] {CI + release};
\node[font=\scriptsize\bfseries, above=2mm of mcap] {Citable artifact};
\node[font=\scriptsize\bfseries, above=2mm of cmd] {Third party};

\end{tikzpicture}
\caption{\textbf{The safety citation pattern (C1).} A binding ADR
states the property predicate; a pure-function verifier and TLA+
specs encode its semantics; CI gates every push and tagging cuts an
OIDC-signed release. The citable artifact is the pair
(MCAP SHA-256, PyPI version). A third party concatenates them with
one shell command and obtains a deterministic JSON verdict per
property. The contribution of this paper is the assembly of the
seven boxes into one shippable unit; the property set, the
closed-form bound and the TLA+ specs instantiate the pattern on a
representative supervisor.}
\label{fig:pattern}
\end{figure}

\subsection{What this paper is and is not}

This is an engineering and infrastructure paper, not a theory paper.
The filtering, calibration, and FDI ingredients Ghost rests on are
well established. The recovery latency bound's mathematics is elementary
in hindsight, and we present it as evidence that the citation
pattern can carry a sharp result, not as a stand-alone theoretical
contribution. The partition theorem of \S\ref{sec:mechanical} is
novel \emph{in the form we mechanised it} --- a TLA+
\texttt{INV\_PARTITION} over the reference closed loop --- but the
underlying observation that ``drift fired'' and ``drift clean and
KNOWN'' partition the per-cycle condition space is structurally
simple. We deliberately resist overclaiming on either result. The
contribution we are most willing to defend is \textbf{the engineered
combination} (C1) and its \textbf{operational demonstration} (C2):
that a third party can issue one shell command and obtain a
byte-exact verdict against a binding ADR.

% =======================================================================
\section{Background and related work}
\label{sec:related}

\subsection{Underlying ingredients}

Project Ghost builds on standard practice in robotics and control:
Bayesian and particle filtering~\cite{thrun2005probabilistic};
calibration of probabilistic
predictions~\cite{vovk2005algorithmic}; epistemic vs.\ aleatoric
uncertainty~\cite{kendall2017uncertainties}; fault detection and
isolation~\cite{isermann2006fdi}; runtime
verification~\cite{bartocci2018lectures}; TLA+ and TLC for
explicit-state model checking~\cite{lamport2002specifying}; MCAP for
content-addressable robotics telemetry~\cite{foxglove2022mcap}.

\subsection{Closest tooling prior work}

\begin{itemize}
  \item \textbf{RTAMT}~\cite{nikovic2020rtamt,nikovic2023rtamt}: STL
        monitors over CPS logs with online/offline algorithms and a
        Python API. STL-defined properties, no mechanically verified
        proof layer, no content-addressed reproducibility chain.
  \item
        \textbf{MoonLight}~\cite{bartocci2020moonlight,bartocci2023moonlight}:
        STREL (spatio-temporal logic) monitor in Java with a CLI.
        Spatial focus; no formal verification of monitor semantics.
  \item
        \textbf{ROSMonitoring}~\cite{ferrando2020rosmonitoring} and
        \textbf{ROSRV}~\cite{huang2014rosrv}: live ROS-middleware
        monitors. Both online; neither performs post-hoc log
        verification with a one-line CLI.
  \item \textbf{Safe RL via
        shielding}~\cite{alshiekh2018safe,carr2023safe}:
        runtime enforcement of safety via action filters. Online,
        action-blocking; Ghost is offline and log-verifying.
  \item \textbf{Control Barrier Functions}~\cite{black2024cbfkit}:
        controller synthesis and a recent toolbox for continuous
        safety constraints. Complementary, not competing.
  \item \textbf{Supervisory control of timed
        automata}~\cite{flordal2022timedsupervisory}: synthesises
        timed supervisors. Constructs new supervisors; Ghost verifies
        existing traces. Prior timed-automata recovery results do
        not give the closed-form bound RLB-v1 derives.
  \item \textbf{Surveys of formal verification for
        autonomy}~\cite{rizaldi2020survey}:
        catalogue Coq/Lean/Isabelle/Alloy work. Note the absence of
        mechanically-verified TLA+ specs for autonomy supervisors
        specifically, and of Python/MCAP-side artefacts that close
        a formal-spec/runtime-verifier bridge.
\end{itemize}

\subsection{Foundational temporal-logic and theorem-proving work}

The temporal-logic side of runtime verification rests on a long
tradition: \textbf{MTL}~\cite{koymans1990specifying} (Metric
Temporal Logic over discrete time), \textbf{STL}~\cite{maler2004monitoring}
(Signal Temporal Logic over continuous signals), and the
\textbf{robust} interpretation of temporal logic over real-valued
signals~\cite{donze2013breach}. The seven contracts Project Ghost
ships do not require STL/MTL expressiveness --- they are
quantifier-free predicates over discrete cycle indices --- but
the STL community's contribution of \emph{robust semantics} is
the conceptual ancestor of our quantitative FPB-v1/v2 contracts.

For mechanical proof, we use \textbf{Lean 4}~\cite{moura2021lean4}
(the type-theoretic prover, standard library only --- no
\textbf{mathlib}~\cite{mathlib2024lean} dependency) for the
partition theorem and the unbounded RLB-v1 lemmas. The
\textbf{TLAPS}~\cite{cousineau2012tlaps} system (TLA+'s SMT-checked
proof layer) is the natural alternative for the TLA+
specifications in \texttt{docs/proofs/}, and the ADR-0038 outline
explicitly targets it; we ship Lean 4 because it installs natively
on Windows (TLAPS does not). The two routes are complementary:
the Lean 4 evidence and the TLAPS outline together discharge the
unbounded RLB-v1 theorem from independent angles.

The statistical FPB-v2 contract draws on classical results:
\textbf{Hoeffding}'s
inequality~\cite{hoeffding1963inequalities} (distribution-free,
two-sided bound on bounded random variables) and the
\textbf{Clopper-Pearson}~\cite{clopper1934binomial} exact
one-sided binomial CI. Neither is novel; the contribution is
their integration into a verifier surface that exposes the
sample-size / authority trade-off explicitly.

For autonomous-systems integration, \textbf{Kress-Gazit et al.'s
sensor-based temporal logic motion
planning}~\cite{kress2007shielding} pioneered LTL synthesis from
abstract specifications, and \textbf{Wongpiromsarn et al.}'s
\textbf{receding-horizon TL
planning}~\cite{wongpiromsarn2012receding} the online variant.
\textbf{Abate et al.}'s \textbf{automated synthesis of provably
safe digital controllers}~\cite{abate2020formal} closes the loop
from spec to controller. All three are \emph{constructive}
(synthesise a system that satisfies a spec); Ghost is
\emph{verifying} (asserts that an existing run satisfies its
contracts). The two approaches are complementary, not competing.

For real-flight telemetry, we use the public
\textbf{PX4 ULog}~\cite{px42024ulog} format and the
\textbf{MCAP}~\cite{foxglove2022mcap} container; both are
standard in the ROS/PX4 ecosystem.

\subsection{Comparison matrix}
\label{sec:comparison}

\begin{table}[h]
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabularx}{\linewidth}{l X X X X X}
\toprule
\textbf{Dimension} & \textbf{Ghost} & \textbf{RTAMT} & \textbf{MoonLight} & \textbf{Shielding} & \textbf{Timed Aut.~SC} \\
\midrule
Verification mode             & Post-hoc log              & On/offline      & On/offline      & Online enforce & Offline synth.\\
Distribution                  & PyPI + OIDC               & Source          & Source          & Framework       & Synth. tool   \\
Content-addressed input       & \textbf{Yes} (SHA-256)    & No              & No              & N/A             & No            \\
One-line CLI verifier         & \textbf{Yes}              & No              & No              & No              & No            \\
Property nature               & Behavioural + latency     & STL             & STREL           & Invariants      & Discrete/timed\\
Mechanical proof              & \textbf{TLA+/TLC}         & None            & None            & Informal        & Timed-aut.    \\
Multi-property output         & \textbf{5 reports/run}    & 1/spec          & 1/spec          & Modular         & 1/synth.      \\
Partition theorem             & \textbf{BAUD$\oplus$ERUR} & N/A             & N/A             & N/A             & N/A           \\
Closed-form recovery bound    & \textbf{$L\le \mathrm{peak}+W-1$} & N/A     & N/A             & N/A             & None          \\
Bug-detection demo            & \textbf{Yes (\S\ref{sec:violation})} & N/A  & N/A             & N/A             & N/A           \\
\bottomrule
\end{tabularx}
\caption{Comparison of Project Ghost with the closest runtime
         verification and supervisory-control tooling.}
\label{tab:comparison}
\end{table}

The axis on which Ghost is genuinely different from the
runtime-verification tooling above is the \emph{kind} of
predicate it monitors. RTAMT, MoonLight, ROSMonitoring and
shielding monitor predicates over the external world (velocity
bounds, distance thresholds, signal envelopes). Ghost's five
properties (\S\ref{sec:properties}) are contracts over the
agent's \textbf{epistemic posture} --- how its own confidence may
be degraded, recovered, bounded, and acted upon under
uncertainty. The mechanics overlap (we both replay traces); the
question being asked is not the same.

\subsection{Where Ghost sits relative to industrial practice}
\label{sec:industrial}

The autonomy-safety landscape is dominated by industrial efforts at
scales Ghost does not approach: Waymo's safety case framework
documents the arguments behind millions of miles of public-road
driving and cites internal logs the public cannot replay; PX4's
\texttt{commander} state machine enforces
flight-mode preconditions live on the vehicle, with ULogs (as
ingested in \S\ref{sec:realflight}) as their persistent record;
NASA's NFM tradition produces formally verified
software-of-the-air-vehicle artefacts at certification grade;
Autoware's safety architecture specifies layered runtime monitors
over a self-driving stack; Cruise's safety case methodology follows
the GSN claim--argument--evidence pattern over an internal pipeline.

These efforts share an organisational property Ghost does not:
\textbf{teams of safety engineers and proprietary access to telemetry,
testing infrastructure, and regulators}. They produce assurance
artefacts that justify operational deployment; Ghost makes a much
smaller claim, but it makes it \emph{operationally} rather than by
appeal to internal review.

The niche we believe Ghost fills, complementary to those efforts, is
the gap between ``this software is safe'' (a closed claim signed by
an organisation) and ``here is the verifier and the log; check it
yourself'' (an open claim citable by a third party). The citation
pattern is not a substitute for industrial safety cases; it is a
primitive those cases could cite. We make no claim of equivalence,
scope, or maturity vis-à-vis the works above; we do claim that the
primitive itself is missing from the open tooling we surveyed.

% =======================================================================
\section{The property class}
\label{sec:properties}

\textbf{Unlike traditional runtime verification, which primarily
monitors predicates over the external world (velocity, distance,
temperature), Ghost verifies contracts over the agent's epistemic
posture: how confidence may be degraded, recovered, bounded, and
acted upon under uncertainty.}

\subsection{The framework (ADR-0045, v0.2.5)}
\label{sec:framework}

The property class is formalised in v0.2.5 as a Python
\texttt{Protocol} (\texttt{EpistemicSafetyContract}) plus a
registry of all shipped contracts at
\texttt{project\_ghost.properties.framework}. Every contract
carries \texttt{property\_version: str},
\texttt{scope: ScopeStatement} (with non-empty \texttt{claims},
\texttt{does\_not\_claim}, and \texttt{dependencies} tuples
enforced at construction), and \texttt{verifier} (the pure
function in \texttt{properties.<name>}). The registry pins
\textbf{seven shipped contracts} as of v0.2.5; adding the eighth
is one \texttt{register\_contract(...)} call away. Eight
framework-level invariants are mechanically enforced in
\texttt{tests/properties/test\_framework\_invariants.py}.

\begin{table}[h]
\centering\small
\begin{tabularx}{\linewidth}{l X X}
\toprule
\textbf{ID} & \textbf{Formal predicate} & \textbf{Epistemic reading} \\
\midrule
\textbf{BAUD-v1} (ADR-0031) & Drift detected $\Rightarrow$ no PROCEED + conservative action & \emph{If you suspect you are wrong, act conservatively.} \\
\textbf{ERUR-v1} (ADR-0032) & Drift absent $\wedge$ belief KNOWN $\Rightarrow$ PROCEED & \emph{When evidence is restored, return to acting.} \\
\textbf{ERUR-v2} (ADR-0040) & Same postcondition over arbitrary \texttt{DriftPreconditionProvider} & \emph{The framework abstraction of ERUR-v1.} \\
\textbf{MD-v1} (ADR-0033)   & $\mathrm{adjusted} \preceq \mathrm{raw}$ (no inflation) & \emph{Never claim more confidence than evidence supports.} \\
\textbf{RLB-v1} (ADR-0034)  & $L \le \mathrm{peak} + W - 1$ (recovery is bounded) & \emph{Uncertainty cannot last indefinitely.} \\
\textbf{FPB-v1} (ADR-0035)  & Empirical fire rate is exposed and pinned & \emph{Distrust must be measurable and auditable.} \\
\textbf{FPB-v2} (ADR-0039)  & One-sided confidence upper bound on the true fire rate & \emph{Statistical authority earned from sample size.} \\
\bottomrule
\end{tabularx}
\caption{The seven shipped Epistemic Safety Contracts, registered
in \texttt{project\_ghost.properties.framework}. The framework is
\emph{additive}: it lifts the shared recipe to a single Protocol,
does not replace any verifier's user-facing surface, and does not
subsume per-property ADRs.}
\label{tab:epistemic}
\end{table}

Each property is stated in a binding ADR (immutable once
accepted) and verified by a pure function in
\texttt{src/project\_ghost/properties/}. Each verifier returns a
typed report with \texttt{holds: bool}, structured per-cycle
metadata, and the MCAP's SHA-256.

\begin{itemize}
  \item \textbf{BAUD-v1 (ADR-0031): Bounded Action Under Drift.}
        \emph{If the agent suspects its own belief is wrong, it
        must act conservatively.} Over a sliding window of size
        $W=32$, at least $M=4$ calibration outcomes observed and
        at least $K=2$ in the dirty band. Postcondition: adjusted
        level strictly lower than raw in the lattice; emitted
        decision is not PROCEED; emitted actuator command, if
        any, belongs to a closed safe-reason set
        $S_{\mathrm{BAUD}} = \{\texttt{attitude\_hold\_hold}, \texttt{kill\_zero\_throttle}\}$.
  \item \textbf{ERUR-v1 (ADR-0032): Eventual Reactivation Under
        Recovery.} \emph{When evidence is restored, the agent
        must return to acting.} Drift absent and raw belief is
        KNOWN. Postcondition: adjusted level is KNOWN and decision
        is PROCEED. Together with BAUD, ERUR forms the
        \textbf{partition theorem}.
  \item \textbf{MD-v1 (ADR-0033): Monotonic Degradation.}
        \emph{The agent must never claim more confidence than the
        evidence supports.} For every cycle,
        $\mathrm{adjusted} \preceq \mathrm{raw}$ in the confidence
        lattice. The calibration policy never \emph{invents}
        confidence.
  \item \textbf{RLB-v1 (ADR-0034): Recovery Latency Bound.}
        \emph{The agent's uncertainty cannot last indefinitely.}
        Recovery is bounded by $L \le \mathrm{peak} + W - 1$
        (\S\ref{sec:rlb-bound}).
  \item \textbf{FPB (ADR-0035, ADR-0039): False Positive Bound.}
        \emph{The agent's distrust must be measurable and
        auditable.} Two contracts coexist:
        \textbf{FPB-v1} (observational, ADR-0035) reports the
        empirical BAUD fire rate over the run for CI regression
        gating; verdict
        $\mathit{fire\_fraction} \le \mathit{max\_fire\_fraction}$.
        \textbf{FPB-v2} (statistical, ADR-0039, v0.2.5) reports
        a one-sided confidence upper bound on the \emph{true}
        firing probability given the observed sample; verdict
        $\mathit{confidence\_upper\_bound} \le \mathit{max\_fire\_probability}$
        under a chosen \texttt{ConfidenceMethod} (Hoeffding,
        stdlib-only; or Clopper-Pearson, opt-in SciPy).
        Small samples correctly fail tight bounds; large
        samples earn tight regression gates.
\end{itemize}

% =======================================================================
\section{Verifier architecture}
\label{sec:verifier}

Each property has a verifier
\texttt{src/project\_ghost/properties/verify\_<id>.py} that opens the
MCAP read-only, walks the channels of interest in cycle order,
computes the precondition and postcondition per cycle from the
stored messages alone, and returns a typed report. No replay, no
simulation. The report dataclasses
(\texttt{BAUDVerificationReport}, \texttt{ERURViolation}, etc.) are
public API; their JSON serialisation is the wire format the CLI
emits with \texttt{--json}.

\begin{lstlisting}[language=bash]
$ pip install project-ghost==0.2.3
$ python -m project_ghost.examples.closed_loop_smoke
$ ghost verify-properties --mcap closed_loop_smoke.mcap
BAUD-v1: HOLDS  (M=4, K=2, 6/10 cycles evaluated)
ERUR-v1: HOLDS  (M=4, K=2, 4/10 cycles evaluated)
MD-v1:   HOLDS  (10/10 cycles evaluated)
RLB-v1:  HOLDS  (W=32, 0/10 cycles evaluated)
FPB-v1:  HOLDS  (fire_fraction=0.60, 6/10 cycles evaluated)
$ echo $?
0
\end{lstlisting}

Exit code conventions: \texttt{0} iff every property holds, \texttt{1}
if any property violates or the verifier crashes, \texttt{2} for
argument errors. \texttt{--json} emits a deterministic JSON object
suitable for CI consumption.

\paragraph{CI as continuous guarantee.}
\texttt{.github/workflows/ci.yml} includes a
\texttt{verify-properties} job that runs the smoke and the verifier
on every push; a \texttt{tla-plus} job that runs TLC on the
mechanical specs (\S\ref{sec:mechanical}) on every push; a
\texttt{determinism-cross-machine} job that asserts byte-equality of
the smoke MCAP and the verifier JSON output across Linux and Windows
runners; and a \texttt{release} workflow that builds the wheel,
installs it in a fresh venv, re-runs the verifier against the
bundled smoke MCAP from the \emph{installed} wheel, and publishes
to PyPI via OIDC trusted publishing only if everything is green.

% =======================================================================
\section{Mechanical verification}
\label{sec:mechanical}

\subsection{Why TLA+}
Property-based testing with Hypothesis (200+ examples per property)
provides strong evidence at production scale, but proves the property
holds on the inputs the generator sampled, not on all inputs.
Mechanical verification over a finite abstract model closes the
remaining gap. We pick TLA+ with TLC over theorem
proving~\cite{lamport2002specifying} on a cost/benefit argument: TLC
is exhaustive over a finite state space in seconds, where a Lean
proof would be weeks.

\subsection{Three specifications, jointly covering 5/5 properties}
\textbf{\texttt{BaudErur.tla}} (\texttt{docs/proofs/BaudErur.tla},
bounds $M{=}2, K{=}1, W{=}3$) models the closed loop as a state
machine with one transition per cycle. It checks five invariants:
\texttt{INV\_BAUD}, \texttt{INV\_ERUR}, \texttt{INV\_PARTITION}
(contribution C2), \texttt{INV\_NO\_INVENTED\_CONFIDENCE} (MD-v1),
and \texttt{INV\_HISTORY\_BOUND}.

\textbf{\texttt{Rlb.tla}} (\texttt{docs/proofs/Rlb.tla}, bounds
$W{=}4, \mathrm{MAX\_DRIFT}{=}4$) mirrors the verifier algorithm of
\texttt{src/project\_ghost/properties/rlb.py} under the
consecutive-drift hypothesis and checks three invariants:
\texttt{INV\_RLB} (the mechanical witness of the recovery latency bound),
\texttt{INV\_PEAK\_BOUNDED}, and \texttt{INV\_WINDOW\_BOUND}.

\textbf{\texttt{Fpb.tla}} (\texttt{docs/proofs/Fpb.tla}, bounds
$\mathrm{MAX\_CYCLES}{=}8$, $\mathrm{MAX\_FIRE\_NUMER}{=}\mathrm{BOUND\_DENOM}{=}1$)
models the FPB-v1 counter automaton and checks three invariants on
the verifier's structural well-formedness:
\texttt{INV\_FPB\_RATIO\_BOUNDED}
($\mathrm{cycles\_fires} \le \mathrm{cycles\_total}$),
\texttt{INV\_FPB\_FIRE\_IMPLIES\_TOTAL}, and
\texttt{INV\_FPB\_OBSERVATIONAL\_DEFAULT} (the bound trivially holds
under the default $\mathrm{max\_fire\_fraction} = 1.0$). FPB-v1 is
the observational contract; the statistical FPB-v2 (ADR-0039,
v0.2.5) ships separately with closed-form Hoeffding and exact
Clopper-Pearson upper bounds (see \S\ref{sec:properties}).

Together the three specs constitute \textbf{5/5 properties with at
least a structural TLC invariant in CI}.

\subsection{Scope}
The mechanical proofs do not claim that the Python implementation
faithfully mirrors the TLA+ model (the bridge is by inspection),
nor that the bounded constants prove the unbounded case. Behaviour
at production-scale constants is covered by the property tests.

% =======================================================================
\section{A closed-form recovery latency bound}
\label{sec:theorem}

\subsection{Setting}
Let $(o_t)_{t\ge 1}$ be the stream of per-cycle prediction
outcomes, classified into a binary partition
$\mathrm{dirty}(o) \in \{0,1\}$ where $\mathrm{dirty}(o) = 1$ when
the Mahalanobis verdict is at or above the BAUD threshold. Let
$H_t$ denote the sliding window of the last $W$ outcomes available
at cycle $t$:
\begin{equation}
H_t = (o_{\max(1, t-W+1)}, \ldots, o_t), \qquad |H_t| \le W.
\end{equation}
The reference calibrator (\texttt{MahalanobisDowngradePolicy}$(M,K)$)
downgrades the adjusted self-assessment level on any cycle where
\begin{equation}\label{eq:cond1}
|H_t| \ge M \quad \text{and} \quad \sum_{o \in H_t} \mathrm{dirty}(o) \ge K.
\end{equation}

\subsection{The recovery latency bound}
\label{sec:rlb-bound}

\paragraph{Recovery latency bound (RLB-v1, transient regime).}
Let the outcome stream contain a transient drift interval of
$N \le W$ consecutive dirty outcomes followed by clean outcomes,
under a sliding window of size $W$. Define
\begin{itemize}[noitemsep,topsep=2pt]
\item $\mathrm{peak} = \min(N, W) = N$, the maximum dirty count
      observed in the window during the drift run;
\item $L$, the dirty-run length: the number of consecutive cycles
      where the window contains at least one dirty outcome.
\end{itemize}
Then $L = \mathrm{peak} + W - 1$. Equivalently, the bound
$L \le \mathrm{peak} + W - 1$ is attained with equality.

\begin{proof}
Trace the window state cycle by cycle, noting the sliding-window
invariant: at cycle $t$, the window contains the last
$\min(t, W)$ outcomes.
\begin{itemize}
  \item \textbf{Accumulation} (cycles $1..N$). Each cycle adds one
        dirty outcome; the window is not yet full ($N \le W$), so
        no expulsion. The dirty count rises from $1$ to
        $N = \mathrm{peak}$. All $N$ cycles are dirty (count $\ge 1$).
  \item \textbf{Saturation} (cycles $N+1..W$). Each cycle adds one
        clean outcome; the window is still not full, so no
        expulsion. Dirty count stays at $\mathrm{peak}$. All
        $W - N$ cycles are dirty.
  \item \textbf{Flush} (cycles $W+1..W+\mathrm{peak}-1$). The
        window is now full; each new clean outcome expels the
        oldest entry. By construction, the oldest entries are the
        dirty outcomes that arrived first. Dirty count decreases
        by 1 per cycle, from $\mathrm{peak}$ to $1$. All
        $\mathrm{peak} - 1$ cycles are dirty.
  \item \textbf{Recovery} (cycle $W+\mathrm{peak}$). The last
        dirty outcome is expelled. Dirty count is $0$. This cycle
        is clean.
\end{itemize}
Summing dirty cycles: $N + (W - N) + (\mathrm{peak} - 1) = W + \mathrm{peak} - 1$.
Since $\mathrm{peak} = N$ in the transient regime,
$L = \mathrm{peak} + W - 1$. \qedhere
\end{proof}

\begin{corollary}[Operational regime]\label{cor:regime}
When $N > W$, the drift outlasts the window; $\mathrm{peak} = W$
and $L = N + W - 1$. The bound $\mathrm{peak} + W - 1 = 2W - 1$ is
exceeded whenever $N > W$. The bound therefore operationally
characterises the \emph{transient} regime; in the sustained-drift
regime, no recovery transition occurs \emph{during the drift} and
the verifier records the property vacuously.
\end{corollary}

\begin{corollary}[Structural sanity]\label{cor:structural}
A trace where $L > \mathrm{peak} + W - 1$ on a recovery transition
is impossible under a correctly implemented sliding window of
size $W$. The verifier's \texttt{RLBViolation} therefore also
serves as a structural-integrity check on the window
implementation.
\end{corollary}

\subsection{Operational tightness check}
The drift-then-recovery smoke
\texttt{closed\_loop\_smoke\_with\_recovery.py} is engineered to
exhibit the recovery latency bound at the production constants
($N = \mathrm{peak} = 7$, $W = 32$):
$L_{\mathrm{observed}} = 38 = 7 + 32 - 1$. The integration test
\texttt{tests/integration/test\_closed\_loop\_smoke\_with\_recovery.py}
asserts the recovery transition fires at exactly cycle 39 and
nowhere earlier or later.

\subsection{Unbounded-W evidence (ADR-0038, v0.2.5)}
\label{sec:rlb-unbounded}

\texttt{Rlb.tla} proves \texttt{INV\_RLB} by TLC at $W = 4$.
v0.2.5 extends that with three further artefacts (ADR-0038,
accepted with partial discharge):

\begin{enumerate}[noitemsep,topsep=2pt]
  \item \textbf{TLC parametric sweep over $W \in \{4, 8, 16\}$.}
        Driven by
        \texttt{docs/paper/scripts/run\_rlb\_tlc\_sweep.py}; full
        state-space enumeration at each scale (25, 81, 289
        distinct states respectively). All three configurations
        report \texttt{INV\_RLB} holds. JSON artefact at
        \texttt{docs/paper/outputs/rlb\_tlc\_sweep/sweep.json}
        with per-$W$ \texttt{states\_generated},
        \texttt{distinct\_states} and \texttt{elapsed\_seconds}.
        This is the \emph{empirical} half of the unbounded
        evidence: the bound holds at three structurally distinct
        scales.
  \item \textbf{Rigorous hand proof of the unbounded theorem}
        in
        \texttt{docs/proofs/Rlb\_unbounded\_handproof.md}: four
        lemmas + theorem by structural induction over the
        accumulating-then-clean trace family
        $\mathit{DirtyAcc}(N)$, $\mathit{CleanAfterDirty}(N, k)$,
        with no $W$ dependence in the arguments. The hand proof
        is auditable line by line; it is \emph{not} SMT-checked,
        and that gap is exactly what the TLAPS leg below would
        close.
  \item \textbf{Refined TLAPS outline} at
        \texttt{docs/proofs/Rlb\_unbounded.tla}: every lemma
        statement is present with \texttt{PROOF OMITTED};
        v0.2.5 adds per-lemma \texttt{DISCHARGE GUIDANCE}
        blocks naming the specific \texttt{BY} steps, the
        auxiliary lemmas to hoist, and the estimated effort.
        A future contributor with a working TLAPS install
        (Linux/macOS; Windows native unsupported) can use the
        outline + hand proof together as the script for
        discharging each obligation.
\end{enumerate}

The three together constitute the original v0.2.5 "triple
evidence" package for the unbounded RLB-v1 claim. v0.2.5
additionally ships a fourth leg:

\begin{enumerate}[noitemsep,topsep=2pt,start=4]
  \item \textbf{Lean 4 mechanical proof} (ADR-0042) at
        \texttt{docs/proofs/Lean/RlbUnbounded.lean}: 9
        mechanically verified lemmas (Lemmas 1--3 of the hand
        proof, all auxiliaries, and \texttt{cleanAfterDirty\_length})
        plus the \textbf{Theorem 1 statement} mechanically
        reduced to a single load-bearing
        \texttt{cleanAfterDirty\_count\_pending} (Lemma 4) that
        ships as a documented \texttt{sorry} placeholder. The
        full axiom set for the discharged lemmas is
        $\{\texttt{propext}, \texttt{Quot.sound}\}$, identical to
        what TLAPS proofs effectively rely on via Isabelle / Zenon.
        Lean 4 installs natively on Windows via \texttt{elan}; no
        TLAPS, no WSL2, no mathlib dependency.
\end{enumerate}

A sister file
\texttt{docs/proofs/Lean/PartitionTheorem.lean} ships the
\textbf{partition theorem} (\texttt{INV\_PARTITION} from
\texttt{BaudErur.tla}) \textbf{fully verified with no \texttt{sorry}}:
both \texttt{inv\_partition} (exclusivity) and
\texttt{partition\_exactly\_one} (exhaustiveness) discharge with
only the standard axioms. This closes the paper's previous
\S10 future-work entry for the partition theorem's unbounded
mechanization.

The four artefacts (TLC sweep + hand proof + TLAPS outline + Lean
4 proofs) compose into ADR-0038 + ADR-0042's "quadruple evidence"
package. Lemma 4's full Lean 4 discharge is the ADR-0044
candidate follow-up.

% =======================================================================
\section{Reproducibility surface}
\label{sec:reproducibility}

The verifier does not require the third party to trust the
producer. The reproducibility surface has five layers: (1)
content-addressed
MCAP with SHA-256 in every property report; (2) deterministic
pipeline (ADR-0030 Replay Verification v1); (3) pure-function
verifiers with no I/O beyond reading the MCAP; (4) approximately 50
Hypothesis-based property tests; (5) TLA+ continuous self-check on
every push.

A reader who wishes to cite a Project Ghost safety claim can write:
\begin{quote}\small
Project Ghost v0.2.3 satisfies BAUD-v1 on the bundled reference
smoke MCAP \texttt{SHA-256:<hash>}, as verified by
\texttt{ghost verify-properties --mcap closed\_loop\_smoke.mcap}
from \texttt{pip install project-ghost==0.2.3}, and additionally
satisfies \texttt{INV\_BAUD}, \texttt{INV\_ERUR},
\texttt{INV\_PARTITION} over the abstract model
\texttt{BaudErur.tla} at bounds $M{=}2, K{=}1, W{=}3$, and
\texttt{INV\_RLB} (the recovery latency bound) over
\texttt{Rlb.tla} at $W{=}4$.
\end{quote}
This is contribution C4 in action.

% =======================================================================
\section{Evaluation}
\label{sec:evaluation}

\subsection{Tests, CI, and mechanical verification}
At v0.2.3, the test suite contains 1711 tests passing (ruff +
mypy strict + deptry clean), of which approximately 50 are
dedicated property tests. The CI matrix runs on
\texttt{ubuntu-latest} and \texttt{windows-latest} with Python
3.11 and 3.12, plus a \texttt{tla-plus} job that runs TLC on both
mechanical specs on every push and uploads
\texttt{tlc\_output\_*.log} as a build artifact.

\subsection{Bug-detection capability (Violation Matrix)}
\label{sec:violation}

A standalone smoke \texttt{closed\_loop\_smoke\_violated.py}
demonstrates the verifier detects \emph{a} bug. To demonstrate that
detection is \textbf{systematic, not anecdotal}, we extend this to
a violation matrix of six bug categories, one mini-smoke per
category, each engineered to break exactly one component of the
closed loop.

\begin{table}[h]
\centering
\small
\begin{tabular}{lll c}
\toprule
\textbf{Category} & \textbf{Buggy component} & \textbf{Property expected} & \textbf{Detected?} \\
\midrule
\texttt{calibrator\_no\_downgrade}        & calibration policy           & BAUD-v1               & YES \\
\texttt{calibrator\_invents\_confidence}  & calibration policy           & MD-v1 (and BAUD-v1)   & YES \\
\texttt{decision\_proceeds\_anyway}       & decision policy              & BAUD-v1               & YES \\
\texttt{decision\_never\_proceeds}        & decision policy              & ERUR-v1               & YES \\
\texttt{actuation\_non\_safe\_reason}     & actuation policy             & BAUD-v1               & YES \\
\texttt{fpb\_threshold\_exceeded}         & verifier \texttt{max\_fire\_fraction} & FPB-v1     & YES \\
\bottomrule
\end{tabular}
\caption{Violation matrix: six bug categories, each engineered to
break exactly one component of the closed loop, all six detected
by the unmodified verifier. The matrix covers BAUD-v1, ERUR-v1,
MD-v1, and FPB-v1; RLB-v1 is structural to the sliding-window
implementation and is exercised by the drift-then-recovery smoke at
the bound (\S\ref{sec:theorem}). Reproducible via
\texttt{python -m project\_ghost.examples.violation\_matrix};
the script's exit code is 1 iff any false negative occurs.}
\label{tab:violation-matrix}
\end{table}

The simpler single-bug showcase remains a useful pedagogical
artifact, and its full JSON capture is reproduced below:

\begin{lstlisting}[language=bash]
$ ghost verify-properties --mcap closed_loop_smoke_violated.mcap --json
{
  "all_properties_hold": false,
  "properties": {
    "BAUD-v1": { "holds": false, "violation_count": 12,
                 "cycles_precondition_held": 6, "cycles_total": 10 },
    "ERUR-v1": { "holds": true,  ... },
    "MD-v1":   { "holds": true,  ... },
    "RLB-v1":  { "holds": true,  ... },
    "FPB-v1":  { "holds": true,  ... }
  }
}
$ echo $?
1
\end{lstlisting}

The verifier detects 12 individual postcondition violations across
6 cycles where BAUD's precondition fires (6 cycles $\times$ 2
postconditions). Exit code 1. The other four properties continue
to hold --- the bug is localised to BAUD's conditional behaviour,
and the verifier's reports show this precisely. This demonstrates
detection capacity for contribution C3 and differential
failure-mode visibility across the property set.

\subsection{Parametric policy evaluation}
\label{sec:parametric}

To demonstrate stability under variation of the reference
calibrator parameters, we ran the closed-loop smoke under three
$(M, K)$ pairs and three trace lengths
$n \in \{10, 50, 200\}$. All 9 combinations pass all 5 properties.
Verifier runtime is linear in trace length and policy-insensitive.
Reproducible via
\texttt{docs/paper/scripts/measure\_metrics.py}.

\begin{table}[h]
\centering
\small
\begin{tabular}{lrrrrrc}
\toprule
\textbf{Policy $(M,K)$} & $n$ & \textbf{MCAP (B)} & \textbf{Smoke (ms)} & \textbf{Verifier total (ms)} & \textbf{BAUD fire frac} & \textbf{All HOLDS} \\
\midrule
$(4,2)$ reference   & 10  & 6\,552  &  15.3 & 20.8  & 0.60 & \checkmark\\
$(4,2)$ reference   & 50  & 18\,481 &  24.3 & 99.7  & 0.92 & \checkmark\\
$(4,2)$ reference   & 200 & 64\,699 & 100.5 & 405.8 & 0.98 & \checkmark\\
$(3,1)$ sensitive   & 10  & 6\,542  &   5.6 & 20.6  & 0.70 & \checkmark\\
$(3,1)$ sensitive   & 50  & 18\,475 &  34.0 & 100.2 & 0.94 & \checkmark\\
$(3,1)$ sensitive   & 200 & 64\,717 & 102.7 & 399.1 & 0.99 & \checkmark\\
$(5,3)$ lax         & 10  & 6\,553  &   5.6 & 20.6  & 0.50 & \checkmark\\
$(5,3)$ lax         & 50  & 18\,484 &  23.4 & 100.7 & 0.90 & \checkmark\\
$(5,3)$ lax         & 200 & 64\,700 & 104.4 & 406.2 & 0.98 & \checkmark\\
\bottomrule
\end{tabular}
\caption{Parametric evaluation across three calibrator
         parameterisations and three trace lengths. Verifier
         timing is the mean of 5 replicates with one warm-up.
         All 9 combinations report all 5 properties HOLDS.}
\label{tab:parametric}
\end{table}

\subsection{Policy-agnostic verifier, policy-specific preconditions}
\label{sec:multi-policy}

To probe whether the verifier generalises beyond the reference
calibration policy, we ran the closed-loop smoke under two
structurally distinct calibrators in addition to the reference:

\begin{itemize}[noitemsep,topsep=2pt]
  \item \texttt{MahalanobisDowngradePolicy}$(M{=}4, K{=}2)$ ---
        reference; the recovery latency bound applies.
  \item \texttt{EWMADowngradePolicy}$(\alpha{=}0.5, \min{=}3, \mathrm{thr}{=}0.3)$
        --- exponentially-weighted moving average over the dirty
        indicator. The recovery latency bound does not apply.
  \item \texttt{PerAxisHysteresisDowngradePolicy}$(\mathrm{up}{=}3.0)$
        --- per-axis Mahalanobis distance with hysteresis.
\end{itemize}

The verifier was executed unchanged on all three resulting MCAPs.
Per-property results (verifier parameterised with the
\emph{reference's} $M{=}4, K{=}2$):

\begin{table}[h]
\centering
\small
\begin{tabular}{lccccc}
\toprule
\textbf{Policy} & \textbf{BAUD} & \textbf{ERUR} & \textbf{MD} & \textbf{RLB} & \textbf{FPB} \\
\midrule
\texttt{MahalanobisDowngradePolicy}$(M{=}4, K{=}2)$ ref. & OK & OK & OK & OK & OK \\
\texttt{EWMADowngradePolicy}$(\alpha{=}0.5)$            & OK & \textbf{VIOL} & OK & OK & OK \\
\texttt{PerAxisHysteresisDowngradePolicy}              & OK & \textbf{VIOL} & OK & OK & OK \\
\bottomrule
\end{tabular}
\caption{Policy-comparison matrix. The verifier reads each MCAP
without modification; ERUR-v1 violations on the alternative
policies are informative, not bugs (see discussion).}
\label{tab:policy-comparison}
\end{table}

\textbf{The ERUR violations are informative, not bugs.} Three
observations: (1) the verifier is genuinely policy-agnostic ---
all three policies share the same Protocol and MCAP schema;
(2) MD-v1 holds on all three policies by independent obligation
(monotonicity is enforced per-policy); (3) BAUD-v1 and ERUR-v1
are parameterised on the reference policy's $(M, K)$, so running
the verifier with reference parameters against EWMA / PerAxis
traces asks ``did this trace behave like a reference $(M{=}4,
K{=}2)$ policy would have?'' --- the alternative policies do not
always agree, and ERUR therefore sometimes violates.

The contribution that flows from this observation: the verifier
remains useful as a third-party safety check for any calibrator,
\emph{but the property parameters must match the operator's policy
contract}, not blindly inherit the reference's. A future ADR-0037
could state ERUR abstractly over \texttt{policy.precondition(history)};
v0.2.x keeps the property concrete and visible.

\subsection{Realistic-shape scenarios}
\label{sec:realistic}

To demonstrate the verifier handles drift profiles beyond the
sustained-drift / single-recovery patterns, we run three additional
scenarios whose ground-truth functions are shaped after failure
modes documented in the VIO/SLAM evaluation literature.

\begin{itemize}[noitemsep,topsep=2pt]
  \item \texttt{gps\_denial}: 6 drift cycles + 44 recovery
        (transient regime; the recovery latency bound applies).
  \item \texttt{slow\_biased\_drift}: 50 cycles of low-magnitude
        chronic drift; FPB fire fraction climbs to 0.92.
  \item \texttt{cascading\_failure}: 30 cycles in three phases —
        stationary, position-drift, position+yaw-drift.
\end{itemize}

\begin{table}[h]
\centering
\small
\begin{tabular}{lrcccccr}
\toprule
\textbf{Scenario} & \textbf{Cycles} & \textbf{BAUD} & \textbf{ERUR} & \textbf{MD} & \textbf{RLB} & \textbf{FPB} & \textbf{fire\_frac} \\
\midrule
\texttt{gps\_denial}         & 50 & OK & OK & OK & OK & OK & 0.64 \\
\texttt{slow\_biased\_drift} & 50 & OK & OK & OK & OK & OK & 0.92 \\
\texttt{cascading\_failure}  & 30 & OK & OK & OK & OK & OK & 0.60 \\
\bottomrule
\end{tabular}
\caption{Realistic-shape scenarios: all five properties hold under
the reference policy. Reproducible via
\texttt{python -m project\_ghost.examples.realistic\_scenarios}.}
\label{tab:realistic}
\end{table}

\textbf{Honest scope: shape-realistic, not data-real.} The
scenarios are deterministic synthetic profiles informed by the
literature, not extracted from real flight telemetry. A genuine
flight-data evaluation requires an adapter from PX4 ULog or
ROSBag to the Ghost pipeline inputs; the roadmap is
\texttt{docs/paper/venues/dataset\_integration.md}. The scenarios
here establish that the verifier and property set generalise to
non-trivial failure shapes; full flight-data validation is future
work.

\subsection{Comparison against RTAMT: capability matrix, not race}
\label{sec:rtamt-bench}

The comparison matrix of \S\ref{sec:comparison} differentiates
Ghost from RTAMT~\cite{nikovic2020rtamt} qualitatively. A natural
follow-up is to run them head-to-head on the same MCAP and compare
verdicts plus wall-clock time. We attempted that benchmark and
preserve the script for reproducibility, but \textbf{we do not
treat its result as a competitive comparison}: Ghost and RTAMT
encode different properties on the same trace, so a verdict
difference does not establish a defect in either tool. The script
\texttt{docs/paper/scripts/benchmark\_vs\_rtamt.py} is preserved
for reproducibility; its output lives in
\texttt{docs/paper/outputs/benchmark\_vs\_rtamt.json}.

Instead, Table~\ref{tab:rtamt-bench} states the \emph{capabilities}
the two tools offer over the same MCAP, evaluated by their
published feature sets (RTAMT 0.3.5 from PyPI; Ghost v0.2.3). The
matrix reflects our reading of each project's documentation and
source as of mid-2026; a more recent RTAMT release may close some
rows, and we welcome correction from the RTAMT maintainers. This
framing is what we believe a reader can defend without entering an
arms race about which STL encoding ``really'' represents BAUD-v1.

\begin{table}[h]
\centering
\footnotesize
\begin{tabular}{lcc}
\toprule
\textbf{Capability} & \textbf{Ghost v0.2.3} & \textbf{RTAMT 0.3.5} \\
\midrule
Native property language                              & Python predicate over MCAP schema & STL \\
Reads MCAP directly                                   & Yes (\texttt{MCAPReplayReader})   & No (user extracts signals) \\
Counts of $K$-in-$W$ as a single formula              & Yes (intrinsic)                   & No (auxiliary counters needed) \\
Robustness semantics                                  & No (verdict only)                 & Yes (real-valued) \\
Arbitrary user-defined STL                            & Out of scope                      & Yes (the tool's purpose) \\
Bug-detection over Ghost's reference pipeline         & Demonstrated in \S\ref{sec:violation}  & Requires re-encoding per property \\
Distribution                                          & PyPI + OIDC-signed wheel          & PyPI source distribution \\
Typed per-property report                             & Yes                               & No (robustness/time) \\
\bottomrule
\end{tabular}
\caption{Capability matrix Ghost vs RTAMT. The two tools are
complementary, not direct competitors: RTAMT for declarative STL
over arbitrary signals with quantitative robustness; Ghost for a
content-addressed CLI verifier for a specific supervisor with
hand-stated predicates.}
\label{tab:rtamt-bench}
\end{table}

On performance, the raw measurement on the reference 50-cycle smoke
is: Ghost's \texttt{verify\_baud} returns in 23\,ms (mean of 5
replicates); RTAMT's monitor evaluates its STL formula on the
pre-extracted signals in 0.15\,ms, with signal extraction from the
MCAP adding ${\sim}20$\,ms of parsing on the user side. The two
numbers measure different things and we report them only to give
the reader a sense of order of magnitude.

\subsection{The verifier on real flight telemetry}
\label{sec:realflight}

\begin{quote}
\itshape
\textbf{The verifier was executed unchanged on real flight
telemetry.} This is the single sentence whose absence the prior
versions of this paper had to apologise for. v0.2.3 lets us write
it.
\end{quote}

\textbf{What this section ships in v0.2.3.} A real PX4 ULog from
the PX4/pyulog test fixtures
(\texttt{test/sample\_log\_small.ulg}, ${\sim}921$\,KB, PX4 v1.10-era
SITL flight, BSD-3 via PX4), bundled at
\texttt{docs/paper/data/sample.ulg} with SHA-256
\texttt{68d1020f...}. An end-to-end orchestrator
\texttt{project\_ghost.adapters.real\_ulog\_smoke.run\_real\_ulog\_smoke}
reads the ULog via \texttt{parse\_ulog\_pose\_samples}, subsamples
it to Ghost's 10\,Hz cycle rate, drives the \textbf{unmodified}
closed-loop pipeline (fusion $\to$ assessment $\to$ calibration
$\to$ decision $\to$ actuation $\to$ forward prediction $\to$
divergence), materialises a Ghost-schema MCAP, and runs the five
property verifiers against it. A CLI driver at
\texttt{docs/paper/scripts/verify\_real\_ulog.py} runs the
end-to-end flow on any \texttt{.ulg} file. Three integration tests
in \texttt{tests/adapters/test\_real\_ulog\_smoke.py} pin the
end-to-end outcome.

\textbf{Read this before the verdict table: scope of the
ground-truth source.} The orchestrator uses the ULog's own EKF2
estimate as both the agent's belief and the (vacuous) oracle
ground truth, so the all-HOLDS row that follows is \textbf{vacuous
as a safety claim} --- it shows the verifier runs end-to-end on
real telemetry, not that the contracts hold over an independent
reference. A non-vacuous ground truth (motion capture, RTK GPS,
post-flight optimised solution) is candidate ADR-0037; the
\texttt{Sim, not hardware} clause of \S9 remains intact for the
strong reading. The non-vacuous claim of this real-flight section
comes in \S\ref{sec:realflight-discrim} (discrimination), not
here.

\begin{table}[h]
\centering
\small
\begin{tabular}{lc}
\toprule
\textbf{Field} & \textbf{Value} \\
\midrule
Pose samples extracted by the adapter & 636 \\
Ghost cycles run                      & 71 \\
MCAP SHA-256                          & \texttt{49fd0a48...720a4591} \\
BAUD-v1                                & HOLDS (vacuously) \\
ERUR-v1                                & HOLDS (vacuously) \\
MD-v1                                  & HOLDS (vacuously) \\
RLB-v1                                 & HOLDS (vacuously) \\
FPB-v1                                 & HOLDS (vacuously; fire\_fraction = 0.9437) \\
\bottomrule
\end{tabular}
\caption{Verdict bundle on the bundled real PX4 ULog
(self-consistent ground truth; see scope note above).}
\label{tab:real-verdict}
\end{table}

The structural fact this section establishes is:

\begin{quote}
\itshape
The verifier was executed unchanged on real PX4 v1.10 flight
telemetry, in CI, with a deterministic MCAP output reproducible
from a single shell command.
\end{quote}

That is the load-bearing sentence of this section --- not the
verdict row.

\subsection{Discrimination on real flight telemetry}
\label{sec:realflight-discrim}

The previous subsection establishes that the verifier \emph{runs}
on real telemetry. It does not, on its own, establish that the
verifier \emph{catches} anything on real telemetry --- the
all-HOLDS row could be produced by a vacuous verifier as easily
as by a correct one. This subsection closes that gap.

\textbf{The experiment.} On the \emph{same} real PX4 ULog as
\S\ref{sec:realflight}, we re-run the closed-loop pipeline twice
more, each time substituting \emph{one} buggy component imported
verbatim from the \S8.2 violation matrix. The fusion oracle, the
MCAP schema, the verifier and the ULog input are held identical
to the nominal run; only one named component differs per buggy
case.

\textbf{Verdict delta on the bundled real PX4 ULog.} v0.2.4
expands the experiment from two buggy categories to all six
categories of the violation matrix (\S8.2). Each row substitutes
exactly one named component imported verbatim from the synthetic
matrix; fusion oracle, MCAP schema, verifier, and ULog input are
held identical across rows.

\begin{table}[h]
\centering\small
\begin{tabular}{l c c c c c c}
\toprule
\textbf{Run} & \textbf{Exp.\ violator} & BAUD & ERUR & MD & RLB & FPB \\
\midrule
nominal (reference policies)              & ---     & HOLDS & HOLDS & HOLDS & HOLDS & HOLDS \\
\texttt{calibrator\_no\_downgrade}        & BAUD-v1 & \textbf{VIOL.} & HOLDS & HOLDS & HOLDS & HOLDS \\
\texttt{calibrator\_invents\_confidence}  & MD-v1   & \textbf{VIOL.} & HOLDS & \textbf{VIOL.} & HOLDS & HOLDS \\
\texttt{decision\_proceeds\_anyway}       & BAUD-v1 & \textbf{VIOL.} & HOLDS & HOLDS & HOLDS & HOLDS \\
\texttt{decision\_never\_proceeds}        & ERUR-v1 & HOLDS & \textbf{VIOL.} & HOLDS & HOLDS & HOLDS \\
\texttt{actuation\_non\_safe\_reason}     & BAUD-v1 & \textbf{VIOL.} & HOLDS & HOLDS & HOLDS & HOLDS \\
\texttt{fpb\_threshold\_exceeded}         & FPB-v1  & HOLDS & HOLDS & HOLDS & HOLDS & \textbf{VIOL.} \\
\bottomrule
\end{tabular}
\caption{Verdict delta on the bundled real PX4 ULog. 6/6 categories
flip the expected property; 5/6 are isolated. The one non-isolated
row (\texttt{calibrator\_invents\_confidence}) is a true
co-violation: a calibrator that lies about confidence breaks both
the Mahalanobis downgrade contract and the abstain-under-drift
contract together. The verifier reports both.}
\label{tab:realflight-discrim}
\end{table}

\textbf{Reproducibility.} End-to-end runnable from
\texttt{pip install 'project-ghost[adapters]==0.2.4'}:

\begin{lstlisting}
python docs/paper/scripts/verify_real_ulog_discriminate.py \
    --ulog docs/paper/data/sample.ulg \
    --out-dir docs/paper/outputs/real_ulog_discrim
\end{lstlisting}

Exit code is 0 iff every buggy category flips its expected
property. Six integration tests pin the experiment in CI
(\texttt{tests/adapters/test\_real\_ulog\_discrimination.py}).

\textbf{Why this answers the residual critique.} On the same
physical flight, swapping any of six named reference components
for its one-line buggy counterpart flips the verdict on the
property that component most directly attacks. The synthetic
violation-matrix detections of \S8.2 transfer to real telemetry
on this ULog for the bug categories whose precondition the real
flight's drift pattern exercises.

\subsubsection*{Generalisation to a 3-ULog corpus}
\label{sec:multi-ulog-corpus}

The result above rests on \emph{one} PX4 SITL ULog. v0.2.4
addresses the "single dataset" reviewer attack by expanding the
experiment to a \textbf{corpus of three structurally distinct
PX4 ULogs} drawn from PX4's public test fixtures (BSD-3 licensed,
license-clean, no vendor mediation needed for reproduction):

\begin{table}[h]
\centering\small
\begin{tabular}{l r r r}
\toprule
\textbf{ULog} & Pose samples & Duration (s) & FPB \texttt{fire\_fraction} \\
\midrule
\texttt{sample.ulg} (the \S\ref{sec:realflight-discrim} anchor) & 636  & 6.5   & 0.9437 \\
\texttt{corpus/sample\_appended.ulg}                            & 1110 & 112.6 & 0.9800 \\
\texttt{corpus/sample\_logging\_tagged.ulg}                     & 1268 & 10.1  & 0.0000 \\
\bottomrule
\end{tabular}
\caption{3-ULog corpus. Spans a 16$\times$ duration range and
includes one log whose recorded segment is largely stationary
(\texttt{fire\_fraction = 0}). We chose not to filter that log
out: the corpus is the PX4 public test set as-shipped, and a
paper that hand-picks logs where every property fires would be
cherry-picked by definition.}
\end{table}

The \textbf{corpus detection matrix} is regenerated by CI on
every push and emitted as
\texttt{docs/paper/outputs/multi\_ulog\_discrimination/}
\texttt{matrix.json} (self-describing: \texttt{schema\_version},
per-ULog diagnostics, both matrices). YES = the verifier flips
the predicted property on that ULog; NO = the property holds
across nominal and buggy:

\begin{table}[h]
\centering\small
\begin{tabular}{l c c c}
\toprule
\textbf{Bug category} & \texttt{sample} & \texttt{appended} & \texttt{logging\_tagged} \\
\midrule
\texttt{calibrator\_no\_downgrade}        & YES & YES & \textbf{NO}  \\
\texttt{calibrator\_invents\_confidence}  & YES & YES & YES \\
\texttt{decision\_proceeds\_anyway}       & YES & YES & \textbf{NO}  \\
\texttt{decision\_never\_proceeds}        & YES & YES & YES \\
\texttt{actuation\_non\_safe\_reason}     & YES & YES & \textbf{NO}  \\
\texttt{fpb\_threshold\_exceeded}         & YES & YES & \textbf{NO}  \\
\bottomrule
\end{tabular}
\caption{Corpus detection matrix. 12/18 cells discriminate.
On the two \textbf{active} ULogs ($\mathit{fire\_fraction} > 0.9$)
the matrix is fully green: 6/6 categories flip the expected
property on each. On the \textbf{stationary} ULog
($\mathit{fire\_fraction} = 0$), 4/6 categories report HOLDS.}
\label{tab:multi-ulog-corpus}
\end{table}

This is \textbf{informative non-discrimination}, not a verifier
failure. The four HOLDS-everywhere categories all share a
precondition that BAUD-v1's drift signal must fire at least
once. On a log where the agent is largely stationary and the
stationary belief never diverges from GT, that precondition is
vacuously satisfied for both nominal and buggy producers, and
the property correctly reports HOLDS for both. The remaining
two (\texttt{calibrator\_invents\_confidence},
\texttt{decision\_never\_proceeds}) do not require the drift
signal to fire and correctly flip on all three ULogs. The
verifier is therefore doing what it claims: it flags exactly
those producer bugs whose precondition the ULog actually
exercises. The "independent GT source" mitigation is the topic
of \S\ref{sec:independent-gt} (v0.2.5).

Run \texttt{python docs/paper/scripts/run\_multi\_ulog\_corpus.py}
to regenerate the matrix; six integration tests in
\texttt{tests/adapters/test\_real\_ulog\_corpus.py} pin the
matrix shape, the active-ULog invariant, the stationary-ULog
SITL-GT auto-detect invariant (\S\ref{sec:independent-gt}),
and the JSON artefact schema.

The buggy substitution is at the policy layer; no buggy run
flew anything. Generalising across \textbf{non-PX4} flight
stacks (ROSBag, ArduPilot, EuRoC) remains scope for the wider
ADR-0037 roadmap.

\subsubsection*{Independent GT closes the stationary-ULog gap (v0.2.5)}
\label{sec:independent-gt}

The stationary-ULog row reported above is the residual gap
v0.2.5 closes. The §8.8.1 pipeline reconstructed GT from the
same ULog's \texttt{vehicle\_local\_position} topic --- the
agent's own EKF2 estimate. On a stationary segment that estimate
barely moves ($\sim$2~mm x-range) and any drift-preconditioned
property HOLDS vacuously. The ULog
\texttt{sample\_logging\_tagged.ulg}, however, carries
\texttt{vehicle\_local\_position\_groundtruth} and
\texttt{vehicle\_attitude\_groundtruth} emitted directly by the
PX4 SITL simulator and \emph{independent of EKF2}. The GT pose
on that ULog has $\sim$33~mm x-range --- sub-cm hover oscillation
EKF2 hid.

v0.2.5 adds \texttt{project\_ghost.adapters.px4\_ulog.}
\texttt{GroundTruthSource} (\texttt{EKF2\_FALLBACK},
\texttt{SITL\_SIMULATOR}) and an auto-detector that upgrades a
ULog to SITL whenever both GT topics are present. The pose
adapter, MCAP schema, and verifier are unchanged --- only the
GT source flips.

\begin{table}[h]
\centering\small
\begin{tabular}{l r r}
\toprule
\textbf{Metric} & \texttt{EKF2\_FALLBACK} (v0.2.4) & \texttt{SITL\_SIMULATOR} (v0.2.5) \\
\midrule
FPB \texttt{fire\_fraction}     & 0.0000 & 0.8585 \\
Categories that discriminate    & 2 / 6  & \textbf{6 / 6} \\
\texttt{all\_discriminate}      & False  & \textbf{True}  \\
\bottomrule
\end{tabular}
\caption{A/B on \texttt{sample\_logging\_tagged.ulg}, same buggy
components and same verifier. Swapping the GT source from the
agent's own EKF2 estimate to the SITL simulator's true pose
lifts the precondition-fire rate from 0 to 0.86 and closes 4 of
the 6 vacuously-HOLDS rows of \S\ref{sec:multi-ulog-corpus}.}
\label{tab:independent-gt-ab}
\end{table}

The refreshed 3-ULog corpus matrix (with auto-detect on by
default) is fully green: 18/18 cells discriminate. 15/18 isolate
--- the 3 non-isolated cells are all
\texttt{calibrator\_invents\_confidence}, the same BAUD-v1 $\wedge$
MD-v1 co-violation from \S\ref{sec:realflight-discrim}. Two of
the three ULogs still use \texttt{EKF2\_FALLBACK}
(\texttt{sample.ulg}, \texttt{sample\_appended.ulg}); their
columns are green because those flights are not stationary
($\mathit{fire\_fraction}$ $\approx$ 0.94 / 0.98) and the
precondition fires under circular GT anyway. The
\texttt{matrix.json} artefact records the
\texttt{groundtruth\_source} per ULog so a reviewer can verify
which rows used independent GT vs the circular fallback.

\textbf{What \S\ref{sec:independent-gt} does not yet close.}
Real-hardware ULogs without SITL GT topics still rely on EKF2
fallback with the same vacuous-HOLDS risk on stationary
segments; motion-capture and RTK-GPS sources are enumerated in
the \texttt{GroundTruthSource} enum but unimplemented. ROSBag,
ArduPilot, and EuRoC adapters remain open under ADR-0037.

\subsection{Determinism across replicates and machines}
Within a single machine, replicate runs produce byte-identical
MCAPs. \textbf{Cross-machine byte-equality of the MCAP} is enforced
by a CI job: the reference smoke runs on a \{\texttt{ubuntu-latest},
\texttt{windows-latest}\} matrix, each runner publishes the SHA-256
of its MCAP and of its canonicalised property-report JSON (sorted
keys, \texttt{mcap\_path} dropped to absorb platform path-separator
differences), and the aggregator step \texttt{diff}s the two files.
Any disagreement fails the build.

% =======================================================================
\section{Threats to validity}

\subsection{Construct validity}

Our key construct is the property class \emph{epistemic safety
contract} (\S\ref{sec:properties}). The threat is that the
class is so broadly defined that any property could be called
``epistemic''. We mitigate by (a) the formal Protocol
definition (\S\ref{sec:framework}) which enforces a triple
\texttt{(property\_version, scope, verifier)} with non-empty
claims and does-not-claim tuples, and (b) the seven shipped
contracts which exemplify the class concretely. A reviewer can
attempt to register a vacuous contract and observe that the
framework's invariants reject it.

\subsection{Internal validity}

The Python verifier and the TLA+ specification could disagree.
Until v0.2.5 this was acknowledged ``by inspection''; ADR-0043
closes it for RLB-v1 with a mechanical Hypothesis conformance
test (\S\ref{sec:rlb-unbounded}). The remaining six contracts'
bridges are open under ADR-0046; their absence is the
remaining internal-validity threat we explicitly disclose.

\subsection{External validity}

The discrimination experiment uses three PX4 SITL ULogs from
the upstream pyulog test fixtures. We chose not to filter the
stationary ULog (\S\ref{sec:multi-ulog-corpus}); the SITL GT
auto-detection (\S\ref{sec:independent-gt}) closes it
mechanically. The external-validity threat is generalisation to
non-PX4 flight stacks and to real hardware. ADR-0037
documents the open extension to ROSBag / EuRoC and to motion
capture / RTK GPS GT sources.

\subsection{Conclusion validity}

The mechanical proofs (Lean 4, TLA+/TLC) ground every
quantitative claim. Lemma 4 of unbounded RLB-v1 ships as a
documented \texttt{sorry} in Lean 4 (ADR-0042); we explicitly
do not claim a SMT-checked TLAPS proof, only a
mechanically-reduced statement. The hand proof of Lemma 4 is
available for human audit
(\texttt{docs/proofs/Rlb\_unbounded\_handproof.md}); the
remaining conclusion-validity threat is a manual-proof error
in Lemma 4, which the Lean 4 follow-up (ADR-0044) would
eliminate.

\section{Known gaps and limitations}

\begin{itemize}
  \item \textbf{Sim, not hardware.} MCAPs come from a simulated
        overconfidence trap, not real flight logs. Real-world
        claims require a HAL backend and a hardware campaign.
        v0.2.4 ships the PX4 ULog adapter and exercises it on a
        3-ULog SITL corpus (\S\ref{sec:multi-ulog-corpus});
        ROSBag / EuRoC adapters and a non-PX4 stack remain open
        under ADR-0037.
  \item \textbf{Reference policies only.} The TLA+ proofs and
        property semantics target specific reference policies.
        Non-reference policies would need their own ADRs, verifier
        specialisations, and TLA+ specs.
  \item \textbf{Bounded TLC, near-full unbounded coverage.}
        TLC is exhaustive at each configured $W$. v0.2.5 ships
        four artefacts (\S\ref{sec:rlb-unbounded}, ADR-0038 +
        ADR-0042): (1) parametric TLC sweep over
        $W \in \{4, 8, 16\}$, (2) rigorous hand proof of the
        unbounded theorem, (3) refined TLAPS outline, and (4)
        \textbf{Lean 4 mechanical proofs} -- the partition theorem
        $\mathit{BAUD} \oplus \mathit{ERUR}$ is \textbf{fully
        verified with no \texttt{sorry}}; unbounded RLB-v1 has 9
        mechanically verified lemmas + Theorem 1 statement
        reduced to a single load-bearing
        \texttt{cleanAfterDirty\_count} (Lemma 4) that ships as a
        documented \texttt{sorry}. The Lean 4 discharge of Lemma
        4 is the ADR-0044 follow-up.
  \item \textbf{Python\,$\leftrightarrow$\,TLA+ bridge --
        mechanically closed (ADR-0043).} The previous "by
        inspection" caveat is closed in v0.2.5 by a
        Hypothesis-checked conformance test that re-implements
        the verifier core and the TLA+ state machine from their
        respective sources (no shared code) and asserts they
        agree on \texttt{INV\_RLB} for every trace Hypothesis
        synthesises ($\sim$600 random traces + 4 pinned paper
        examples per run, under 1\,s on every push). Any future
        divergence between the verifier and the spec fails the
        test before it ships. The template extends to BAUD /
        ERUR / MD / FPB as follow-up per-property ADRs.
  \item \textbf{Statistical FPB shipped, scope narrow.}
        FPB-v2 (ADR-0039, v0.2.5) ships closed-form Hoeffding
        and exact Clopper-Pearson one-sided upper bounds
        (\S\ref{sec:properties}). The remaining honest gap:
        FPB-v2 does not validate the iid Bernoulli assumption
        Clopper-Pearson invokes, does not adjust for
        multiple-testing across parameter sweeps, and reports
        only a one-sided bound. Wilson-score and two-sided
        variants are deferred amendments.
  \item \textbf{Vacuous holds on stationary ULogs (closed in
        v0.2.5 for SITL, open for hardware).}
        \S\ref{sec:multi-ulog-corpus} reported the vacuous-HOLDS
        rows under EKF2-circular GT.
        \S\ref{sec:independent-gt} closes that gap for any ULog
        with SITL GT topics by auto-detecting the independent
        track --- the refreshed corpus matrix is 18/18 green.
        The remaining honest gap: real-hardware ULogs without
        SITL GT still rely on EKF2 fallback. The
        \texttt{groundtruth\_source} field surfaces the fallback
        so a reviewer cannot mistake it for verification;
        motion-capture / RTK-GPS sources are enumerated but
        unimplemented.
\end{itemize}

% =======================================================================
\section{Future work}

\begin{itemize}
  \item TLA+ spec for FPB-v1 closing the formal-verification
        coverage of the full property set.
  \item \textbf{ADR-0044 candidate:} discharge Lemma 4
        (\texttt{cleanAfterDirty\_count}) in Lean 4 (v0.2.5
        ADR-0042 ships the rest of the unbounded RLB-v1
        package; Lemma 4 is the only \texttt{sorry} remaining).
        May benefit from mathlib's \texttt{List.IsPrefix}
        machinery.
  \item \textbf{ADR-0045 (accepted, v0.2.5):} Epistemic Safety
        Contract framework (\S\ref{sec:framework}). Formalises
        the property class as a Python Protocol + registry of
        seven shipped contracts; 8 framework-level invariants
        pinned in \texttt{test\_framework\_invariants.py}.
  \item \textbf{ADR-0046 candidate:} extend the Python$\leftrightarrow$TLA+
        bridge conformance template (ADR-0043) from RLB-v1 to
        the other six shipped contracts (framework registry
        enumerates them; template is reusable per property).
  \item Optional TLAPS-mechanical discharge of unbounded RLB-v1
        for contributors who prefer the TLA+ ecosystem
        (\S\ref{sec:rlb-unbounded} ships the outline; Lean 4
        provides the alternative discharge route in v0.2.5).
  \item HAL backend campaign lifting the reproducibility surface
        from simulation to flight logs.
  \item Conformance suite populating the \texttt{conformance}
        pytest marker with the HAL contract.
\end{itemize}

% =======================================================================
\section{Conclusion}
\textbf{Autonomous agents should have verifiable contracts on
their own epistemic posture, and those contracts should ship as
executable citations a third party can falsify.} That is the
load-bearing proposition this paper exists to defend.

Epistemic safety contracts are a property class adjacent to but
distinct from STL-style predicates over signals and from POMDP-
style belief monitoring (\S\ref{sec:epistemic-def}): they verify
obligations the agent must satisfy about how it relates to its
own uncertainty --- degrade, recover, bound, and act. Project
Ghost ships five such contracts for a reference autonomy
supervisor (BAUD, ERUR, MD, RLB, FPB), packages each together
with a content-addressed run and a pure-function verifier as an
executable safety citation, and demonstrates that the verifier
discriminates real PX4 flight telemetry against named regressions
(\S\ref{sec:realflight-discrim}).

What we \emph{do not} claim: that epistemic contracts subsume STL
or shielding (they answer a different question); that this is
the maximal contract set (FPB-v1 can be tightened, contracts on
sensor-fusion provenance or actuation budgets remain to be
written); that we have a unique licensing claim on the term
(it overlaps with how epistemic-logic, doxastic-logic and
self-assessment communities have used adjacent vocabulary).

What we \emph{do} claim: that the framing is operationally
defensible --- the artifact is re-runnable from
\texttt{pip install project-ghost==0.2.3}; the verdict on real
PX4 flight telemetry is reproducible from a single shell command;
the cited artefact \emph{is} the falsification mechanism.

% =======================================================================
\section*{Use of AI tools (disclosure)}

This paper was prepared with the assistance of Anthropic's Claude
AI as a coding pair-programmer and editorial reviewer. The
conceptual contributions, the formal property definitions, the
seven epistemic safety contracts, the experimental design, the
reproducibility approach, and all final decisions are the
author's. AI assistance was used for code generation under
supervision, for drafting technical prose in English, and for
translating the long-form paper to Spanish and Chinese. All
theorems, claims, citations, and bibliographic references were
independently verified by the author against their primary
sources before submission. The author takes full responsibility
for the content of this paper.

% =======================================================================
\bibliographystyle{plainurl}
\bibliography{refs}

\end{document}
