% Project Ghost — Workshop / short-paper version (target: 12-15 pages)
% Companion to docs/paper/arxiv/main.tex (the technical-report version).
%
% This short version is the FMAS / RV / SAFECOMP submission target.
% Tables that were in §8 of the long version live in this paper's
% Appendix A; the parametric-eval table moves to Appendix B; the
% RTAMT capability matrix moves to Appendix C. The Main body keeps
% only the comparison matrix (§2.3) and the real-telemetry verdict
% bundle (§7).
%
%   pdflatex main; bibtex main; pdflatex main; pdflatex main

\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{microtype}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[colorlinks=true,linkcolor=blue!60!black,citecolor=blue!50!black,urlcolor=blue!60!black]{hyperref}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,shapes.geometric,backgrounds,fit}

\definecolor{codebg}{rgb}{0.97,0.97,0.97}
\lstset{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{codebg},
  frame=single, rulecolor=\color{black!20},
  numbers=none, breaklines=true,
  showstringspaces=false, columns=fullflexible,
  xleftmargin=0.5em, xrightmargin=0.5em,
}

\newtheorem{theorem}{Theorem}

\title{Epistemic Contracts for Autonomous Systems:\\
       A Verifiable Pattern for Safety Claims Under Uncertainty}
\author{Javier Menéndez Mateos\\
        \texttt{jfhelvetius@gmail.com}\\
        \small Independent}
\date{Short paper, v0.2.3 --- \today}

\begin{document}
\maketitle

\begin{abstract}
\textbf{Thesis: autonomous agents should have verifiable contracts
on their own epistemic posture --- how they degrade confidence,
recover, remain bounded, and translate belief into action under
uncertainty.} Most existing runtime verifiers ask predicates of
the world (velocity, distance, temperature); we propose asking
predicates of the agent's posture toward its own uncertainty. We
call these \textbf{epistemic safety contracts}. We describe
Project Ghost: five contracts for a reference autonomy supervisor
(BAUD/ERUR/MD/RLB/FPB), each verified by a pure function over a
content-addressed MCAP log, with underlying invariants checked
mechanically by TLA+/TLC. Each contract ships together with its
recorded run and verifier as an \textbf{executable safety
citation}: a third party reproduces the verdict from
\texttt{pip install 'project-ghost[adapters]==0.2.3'} followed by
\texttt{ghost verify-properties --mcap <log>}. On real PX4 v1.10
flight telemetry, the verifier \emph{discriminates}: two
independently buggy components, swapped into the same physical
flight, each flip BAUD-v1 from HOLDS to VIOLATED while the four
other contracts remain HOLD.
\end{abstract}

\noindent\textbf{Keywords:} epistemic safety contracts, runtime
verification, uncertainty in autonomy, executable safety
citations, content-addressed telemetry, TLA+, MCAP.

% =======================================================================
\section{Introduction}
\label{sec:intro}

Most existing runtime verifiers ask predicates of the world.
We ask predicates of the agent's posture toward its own
uncertainty. An \textbf{epistemic safety contract} is a triple
$(P,Q,V)$ where $P$ is a predicate over the agent's epistemic
state at cycle $t$, $Q$ a predicate over the agent's behaviour at
$t$, and $V$ a pure-function verifier such that $V(r)$ on a
recorded run $r$ returns HOLDS iff every cycle satisfying $P$
also satisfies $Q$. Distinct from STL (predicates over signals)
and from belief monitoring (a property of what the agent
believes, not of how it must relate to its belief).

We package each contract together with a content-addressed run
and an OIDC-signed verifier wheel as an \textbf{executable safety
citation}: a third party reproduces the verdict from
\texttt{ghost verify-properties --mcap <log>}. \textbf{The cited
artefact \emph{is} the falsification mechanism.}

% =======================================================================
\section{Background and related work}
\label{sec:related}

Ghost rests on standard ingredients: Bayesian filtering
\cite{thrun2005}, calibration \cite{vovk2005}, runtime
verification \cite{bartocci2018}, TLA+/TLC \cite{lamport2002},
MCAP \cite{foxglove2022}. Closest tooling:
RTAMT~\cite{rtamt2020}, MoonLight~\cite{moonlight2020},
ROSMonitoring~\cite{rosmonitoring2020}, ROSRV~\cite{rosrv2014},
shielding~\cite{shielding2020}, control barrier
functions~\cite{cbf2023}, conformal prediction~\cite{lindemann2023},
supervisory control of timed automata~\cite{timedaut2022};
surveys of the field~\cite{rizaldi2020}. None of these ships a
content-addressed pure-function verifier via signed wheels with
a mechanically verified semantics layer; Table~\ref{tab:comparison}
positions Ghost on the relevant axes. Industrial practice (Waymo,
NASA NFM, PX4 \texttt{commander}, Autoware, Cruise) produces
safety artefacts at scales Ghost does not but behind organisational
gates the public cannot replay; Ghost fills the niche of the open,
third-party-citable safety claim. Full per-tool detail in
Appendix~\ref{app:comparison}.

\begin{table}[h]
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabularx}{\linewidth}{l X X X X X}
\toprule
\textbf{Dimension} & \textbf{Ghost} & RTAMT & MoonLight & Shielding & Timed SC \\
\midrule
Mode                  & Post-hoc log         & On/offline    & On/offline & Online & Offline synth.\\
Distribution          & PyPI + OIDC          & Source        & Source     & Framework & Tool \\
Content-addressed in  & \textbf{Yes}         & No            & No         & N/A      & No \\
One-line CLI verifier & \textbf{Yes}         & No            & No         & No       & No \\
Mechanical proof      & \textbf{TLA+/TLC}    & None          & None       & Informal & Timed aut. \\
Multi-property output & \textbf{5/run}       & 1/spec        & 1/spec     & Modular  & 1/synth. \\
Closed-form bound     & $L\le \mathrm{peak}{+}W{-}1$ & N/A     & N/A        & N/A      & None \\
Policy-agnostic?      & \textbf{Yes (v1/v2 lift, App.~C.3)} & STL-specific & STREL-specific & per-controller & per-supervisor \\
Bug-detection demo    & \textbf{Yes (Table~\ref{tab:violation})} & N/A & N/A & N/A      & N/A \\
\bottomrule
\end{tabularx}
\caption{Comparison matrix. No prior tool we surveyed ships a
content-addressed, pure-function safety-property verifier via
\texttt{pip install} with mechanically verified invariants.}
\label{tab:comparison}
\end{table}

The axis on which Ghost is genuinely different is the \emph{kind}
of predicate it monitors. RTAMT, MoonLight and shielding monitor
predicates over the external world; Ghost's five properties
(\S\ref{sec:properties}) are contracts over the agent's
\textbf{epistemic posture}. The \emph{policy-agnostic} row of the
matrix above is realised at the verifier level
(Table~\ref{tab:policy-comparison}, three structurally distinct
calibrators); ERUR is stated in two layers (v1 fixed to the
reference predicate, v2 delegating to the policy's own drift
criterion) so the verifier-agnostic claim does not silently
inherit the reference policy's parameters.

% =======================================================================
\section{The pattern}
\label{sec:pattern}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  font=\footnotesize,
  node distance=4mm and 6mm,
  box/.style={draw, rounded corners=2pt, align=center, inner sep=4pt,
              minimum width=28mm, minimum height=10mm},
  producer/.style={box, fill=blue!8, draw=blue!50!black},
  ci/.style={box, fill=green!10, draw=green!50!black},
  artifact/.style={box, fill=yellow!18, draw=orange!70!black},
  party/.style={box, fill=red!8, draw=red!60!black},
  flow/.style={-{Latex[length=2mm,width=1.6mm]}, thick},
  flowdash/.style={-{Latex[length=2mm,width=1.6mm]}, thick, dashed, gray!70!black},
  flowbig/.style={-{Latex[length=2.6mm,width=2mm]}, ultra thick, draw=red!50!black},
]
\node[producer] (adr) {Binding ADR};
\node[producer, below=of adr] (code) {Pure-function verifier};
\node[producer, below=of code] (spec) {Hypothesis + TLA+};
\node[ci, right=14mm of code] (ci) {CI: verify + TLC + xmach};
\node[ci, below=of ci] (tag) {OIDC-signed wheel};
\node[artifact, right=14mm of ci] (mcap) {MCAP, SHA-256};
\node[artifact, below=of mcap] (cite) {\textbf{The citation}};
\node[party, right=14mm of mcap] (cmd) {\texttt{ghost verify}};
\node[party, below=of cmd] (out) {Verdict + JSON};
\draw[flow] (adr) -- (code);
\draw[flow] (code) -- (spec);
\draw[flow] (spec.east) -- (ci.west);
\draw[flow] (ci) -- (tag);
\draw[flowdash] (code.east) -- ++(3mm,0) |- (mcap.west);
\draw[flow] (tag.east) -- ++(3mm,0) |- (cite.west);
\draw[flow] (mcap) -- (cite);
\draw[flowbig] (cite.east) -- ++(3mm,0) |- (cmd.west);
\draw[flow] (cmd) -- (out);
\end{tikzpicture}
\caption{\textbf{The executable safety citation pattern.} The
citable artefact is the pair (MCAP SHA-256, PyPI version); a third
party reproduces the per-property verdict from one shell command.}
\label{fig:pattern}
\end{figure}

The figure reads left-to-right as the operational pipeline of a
safety claim under the pattern. On the \textbf{producer} side
(blue) a binding ADR states the property predicate, a
pure-function verifier implements its semantics, and Hypothesis
property tests plus TLA+ specifications jointly exercise the
underlying invariants. The \textbf{CI + release} stage (green)
gates every push (verifier + TLC + cross-machine determinism
diff) and, on tagging, emits an OIDC-signed PyPI wheel with no
long-lived token escrowed. The \textbf{citable artefact} (yellow)
carries two halves: the run, materialised as an MCAP log
content-addressed by SHA-256, and the verification tool, pinned by
PyPI version. The \textbf{third party} (red) concatenates the two
halves with one shell command and obtains a deterministic JSON
verdict per property.

The pattern is therefore not just "upload the log next to the
paper". Three structural choices distinguish it from publishing a
PDF + a tarball next to a Zenodo DOI. \emph{First}, the verifier
is part of the citation: a third party does not write their own
oracle, they re-run ours. \emph{Second}, the verifier is a pure
function over a content-addressed log: there is no replay of the
producer, no simulation, no I/O beyond reading the MCAP. The
verifier's verdict therefore depends only on bytes the third party
already holds. \emph{Third}, the underlying invariants the
verifier encodes are themselves mechanically checked by TLA+/TLC
on every push: the third party inherits not only an oracle but the
evidence that the oracle is consistent with the property
statement. The cited artefact is, in this precise sense, the
falsification mechanism.

% =======================================================================
\section{The property set: epistemic contracts}
\label{sec:properties}

\textbf{Unlike traditional runtime verification, which monitors
predicates over the external world (velocity, distance,
temperature), Ghost verifies contracts over the agent's epistemic
posture: how confidence may be degraded, recovered, bounded, and
acted upon under uncertainty.} The five properties are:

\begin{itemize}[noitemsep,topsep=2pt]
  \item \textbf{BAUD-v1.} \emph{If you suspect you are wrong, act
        conservatively.} Drift detected $\Rightarrow$ no PROCEED
        and only a closed safe-reason action set.
  \item \textbf{ERUR-v1 / v2.} \emph{When evidence is restored,
        return to acting.} Two layers: v1 fixes the precondition
        to the reference predicate (\texttt{outcomes < M} or
        \texttt{dirty\_count < K}, with $M=4, K=2$); v2 delegates
        to \texttt{policy.drift\_precondition()}, lifting the
        contract to any calibrator that exposes its own drift
        criterion via the \texttt{DriftPreconditionProvider}
        Protocol. Both verifiers ship in v0.2.4
        (\texttt{verify\_erur} and \texttt{verify\_erur\_v2}); the
        latter accepts a mapping from \texttt{policy\_id} to the
        policy's predicate. Appendix~\ref{app:eval},
        Table~\ref{tab:policy-comparison} compares both verdicts
        on three structurally distinct calibrators (ADR-0040).
  \item \textbf{MD-v1.} \emph{Never claim more confidence than
        evidence supports.} $\mathrm{adjusted} \preceq
        \mathrm{raw}$.
  \item \textbf{RLB-v1.} \emph{Uncertainty cannot last
        indefinitely.} Recovery is bounded by
        $L \le \mathrm{peak} + W - 1$ (\S\ref{sec:mechanical}).
  \item \textbf{FPB (v1 + v2, ADR-0035 / ADR-0039).}
        \emph{Distrust must be measurable and auditable.} v1
        exposes the empirical fire rate as a CI regression gate;
        v2 (v0.2.5) reports a one-sided confidence upper bound
        on the \emph{true} firing probability under Hoeffding
        (stdlib) or Clopper-Pearson (SciPy) --- small samples
        correctly fail tight bounds.
\end{itemize}

Each is verified by a pure function over the content-addressed
MCAP, exercised by Hypothesis property tests, and self-enforced on
every push by CI.

% =======================================================================
\section{Mechanical verification}
\label{sec:mechanical}

Three TLA+ specifications jointly cover the property set with 11
invariants verified by TLC on every push:
\texttt{BaudErur.tla} ($M{=}2, K{=}1, W{=}3$) for BAUD-v1, ERUR-v1,
the partition theorem $\mathrm{BAUD} \oplus \mathrm{ERUR}$, and
MD-v1; \texttt{Rlb.tla} ($W{=}4$) mirrors the verifier algorithm
under the consecutive-drift hypothesis and mechanises the recovery
latency bound $L = \mathrm{peak} + W - 1$ (statement and proof in
\S\ref{app:theorem}); \texttt{Fpb.tla} checks FPB counter
well-formedness.

% =======================================================================
\section{The verifier on real flight telemetry}
\label{sec:realflight}

\textbf{Setup.} A real PX4 ULog from the PX4/pyulog test fixtures
(\texttt{test/sample\_log\_small.ulg}, ${\sim}921$\,KB, PX4 v1.10
SITL flight, BSD-3 via PX4), bundled at
\texttt{docs/paper/data/sample.ulg} with SHA-256
\texttt{68d1020f...}. The end-to-end orchestrator
\texttt{project\_ghost.adapters.real\_ulog\_smoke}
reads it, subsamples to 10\,Hz, drives the unmodified closed-loop
pipeline, materialises a Ghost-schema MCAP, and runs the five
property verifiers.

\subsection{Discrimination, not just all-HOLDS}
\label{sec:realflight-discrim}

A naive run produces all-HOLDS on the real ULog (MCAP SHA-256
\texttt{49fd0a48\ldots}). That row alone does not establish that
the verifier \emph{catches} anything on real telemetry --- a
vacuous verifier would produce the same row. The load-bearing
experiment is the verdict \emph{delta} on the same physical
flight.

\textbf{The experiment.} On the same real PX4 ULog, we re-run the
closed-loop pipeline twice more, each time substituting one buggy
component imported verbatim from the \S2 violation matrix
(Appendix~\ref{app:eval}, Table~\ref{tab:violation}). Fusion
oracle, MCAP schema, verifier and ULog input are held identical to
the nominal run; only one named component differs per buggy case.

\begin{table}[h]
\centering\small
\begin{tabular}{l c c c c c c}
\toprule
\textbf{Run} & \textbf{Exp.\ violator} & BAUD & ERUR & MD & RLB & FPB \\
\midrule
nominal (reference policies)              & ---     & HOLDS & HOLDS & HOLDS & HOLDS & HOLDS \\
\texttt{calibrator\_no\_downgrade}        & BAUD-v1 & \textbf{V.} & HOLDS & HOLDS & HOLDS & HOLDS \\
\texttt{calibrator\_invents\_confidence}  & MD-v1   & \textbf{V.} & HOLDS & \textbf{V.} & HOLDS & HOLDS \\
\texttt{decision\_proceeds\_anyway}       & BAUD-v1 & \textbf{V.} & HOLDS & HOLDS & HOLDS & HOLDS \\
\texttt{decision\_never\_proceeds}        & ERUR-v1 & HOLDS & \textbf{V.} & HOLDS & HOLDS & HOLDS \\
\texttt{actuation\_non\_safe\_reason}     & BAUD-v1 & \textbf{V.} & HOLDS & HOLDS & HOLDS & HOLDS \\
\texttt{fpb\_threshold\_exceeded}         & FPB-v1  & HOLDS & HOLDS & HOLDS & HOLDS & \textbf{V.} \\
\bottomrule
\end{tabular}
\caption{\textbf{Discrimination on real flight telemetry
(v0.2.4).} 6/6 categories flip the expected property; 5/6 are
isolated. The non-isolated row is a true co-violation: an
inflated-confidence calibrator breaks MD-v1 and BAUD-v1 together.}
\label{tab:real-verdict}
\end{table}

\textbf{Corpus generalisation (v0.2.5).} The matrix is exercised
on a 3-ULog PX4 SITL corpus (\texttt{sample.ulg},
\texttt{sample\_appended.ulg}, \texttt{sample\_logging\_tagged.ulg}).
\textbf{18/18 cells discriminate} under the v0.2.5
\texttt{GroundTruthSource} enum with auto-detection: ULogs that
ship \texttt{vehicle\_*\_groundtruth} topics route through an
\emph{independent} SITL GT track instead of the agent's own EKF2
estimate. On the stationary
\texttt{sample\_logging\_tagged.ulg}, switching the GT source
(same buggy components, same verifier) lifts FPB
$\mathit{fire\_fraction}$ from 0 to 0.86 and discriminating
categories from 2/6 to 6/6 --- closing the v0.2.4 vacuous-HOLDS
gap. The
\texttt{docs/paper/outputs/multi\_ulog\_discrimination/}
\texttt{matrix.json} artefact records the
\texttt{groundtruth\_source} per ULog so a reviewer can verify
which rows used independent GT. Real-hardware ULogs without
SITL GT still rely on EKF2 fallback under ADR-0037
(\S\ref{sec:limits}); motion-capture and RTK-GPS sources are
enumerated but unimplemented.

\textbf{Evaluation summary.} Full evaluation in
Appendix~\ref{app:eval}: six-category violation matrix
(Table~\ref{tab:violation}), parametric policy evaluation
(Table~\ref{tab:parametric}), policy-comparison matrix
(Table~\ref{tab:policy-comparison}), realistic-shape scenarios
(Table~\ref{tab:realistic}), RTAMT capability matrix
(Table~\ref{tab:rtamt-bench}).

% =======================================================================
\section{Limitations}
\label{sec:limits}

\begin{itemize}[noitemsep,topsep=2pt]
  \item \textbf{Sim, not hardware.} v0.2.5 ships the PX4 ULog
        adapter, a 3-ULog SITL corpus, and the
        \texttt{GroundTruthSource} enum (SITL auto-detect lifts
        the corpus matrix to 18/18). Real-hardware ULogs without
        SITL GT topics and non-PX4 stacks remain under ADR-0037.
  \item \textbf{Reference policies only.} TLA+ specs target the
        reference. Non-reference policies need their own.
  \item \textbf{Bounded TLC.} Behaviour at production-scale
        constants relies on property tests; TLA+ fills the
        small-but-exhaustive corner.
  \item \textbf{Python$\leftrightarrow$TLA+ bridge by inspection.}
        Automating it is future work.
\end{itemize}

% =======================================================================
\section{Conclusion}
\label{sec:conclusion}

\textbf{Autonomous agents should have verifiable contracts on
their own epistemic posture, and those contracts should ship as
executable citations a third party can falsify.} Epistemic
contracts are adjacent to but distinct from STL-style predicates
over signals and from POMDP-style belief monitoring: they verify
obligations on how the agent must relate to its own uncertainty.
Ghost ships five such contracts for a reference autonomy
supervisor, packages each as an executable safety citation, and
demonstrates discrimination on real PX4 flight telemetry. The
cited artefact \emph{is} the falsification mechanism. The artifact
is re-runnable from
\texttt{pip install 'project-ghost[adapters]==0.2.3'} (the
\texttt{[adapters]} extra brings the PX4 ULog parser used by
\S\ref{sec:realflight}).

% =======================================================================
\bibliographystyle{plainurl}
\bibliography{refs}

% =======================================================================
\appendix

\section{Per-tool comparison detail}
\label{app:comparison}

The runtime-verification literature has produced several tools
with overlapping concerns, none of which occupies the same niche
as Ghost. We summarise each below.

\paragraph{RTAMT} \cite{rtamt2020}.
STL monitors over CPS logs with online/offline algorithms and a
Python API. Property language is signal temporal logic, not
hand-crafted predicates; there is no mechanically verified proof
layer and no content-addressed reproducibility chain.

\paragraph{MoonLight} \cite{moonlight2020}.
STREL (spatio-temporal logic) monitor in Java with a CLI, used
for automotive benchmarks. Spatial focus; no formal verification
of the monitor semantics.

\paragraph{ROSMonitoring} \cite{rosmonitoring2020} \textbf{and ROSRV}
\cite{rosrv2014}.
Live ROS-middleware monitors that intercept the master or wrap
nodes. Both are online; neither performs post-hoc log
verification with a one-line CLI.

\paragraph{Safe RL via shielding} \cite{shielding2020}.
Runtime enforcement of safety via action filters. Online,
action-blocking; Ghost is offline, log-verifying.

\paragraph{Control Barrier Functions} \cite{cbf2023}.
Controller synthesis for continuous safety constraints.
Complementary, not competing.

\paragraph{Conformal prediction for robot safety}
\cite{lindemann2023}.
Forward-looking distribution-free uncertainty bounds for gating
actions. Predictive; Ghost is retrospective.

\paragraph{Supervisory control of timed automata}
\cite{timedaut2022}.
Synthesises timed supervisors. Constructs new supervisors; Ghost
verifies existing traces. Prior timed-automata recovery results
do not give the closed-form bound of \S\ref{app:theorem}.

\paragraph{Surveys of formal verification for autonomy}
\cite{rizaldi2020}.
Catalogue Coq/Lean/Isabelle/Alloy work. Note the absence of
mechanically-verified TLA+ specs for autonomy supervisors
specifically.

To the best of our knowledge, no prior tool ships a
content-addressed, pure-function safety-property verifier via
\texttt{pip install} + OIDC-signed wheels with mechanically
verified underlying invariants. The axis on which Ghost is
genuinely different is the \emph{kind} of predicate it monitors:
RTAMT, MoonLight, ROSMonitoring and shielding monitor predicates
over the external world; Ghost's five contracts
(\S\ref{sec:properties}) are over the agent's epistemic posture.

\section{The recovery latency bound: statement and proof}
\label{app:theorem}

\paragraph{Statement (RLB-v1, transient regime).}
Let the outcome stream contain a transient drift of $N \le W$
consecutive dirty outcomes followed by clean outcomes. Let
$\mathrm{peak} = N$ and $L$ be the dirty-run length. Then
$L = \mathrm{peak} + W - 1$, attained with equality.

\paragraph{Proof.}
Trace the window state cycle by cycle, noting the sliding-window
invariant: at cycle $t$, the window contains the last
$\min(t, W)$ outcomes. Accumulation phase (cycles $1..N$): each
cycle adds one dirty outcome; the window is not yet full
($N \le W$); dirty count rises from $1$ to $N = \mathrm{peak}$.
Saturation phase (cycles $N+1..W$): clean outcomes arrive,
window not yet full, count stays at $\mathrm{peak}$. Flush phase
(cycles $W+1..W+\mathrm{peak}-1$): each clean expels the oldest
(dirty) entry; count drops by 1 per cycle from $\mathrm{peak}$ to
$1$. Recovery at cycle $W+\mathrm{peak}$: last dirty expelled,
count $0$. Total dirty cycles: $N + (W - N) + (\mathrm{peak} - 1)
= W + \mathrm{peak} - 1$. Since $\mathrm{peak} = N$ in the
transient regime, $L = \mathrm{peak} + W - 1$. \qed

\section{Evaluation tables}
\label{app:eval}

\subsection{Violation matrix}

\begin{table}[h]
\centering\small
\begin{tabular}{lll c}
\toprule
\textbf{Category} & \textbf{Buggy component} & \textbf{Property} & \textbf{Detected?} \\
\midrule
\texttt{calibrator\_no\_downgrade}        & calibration policy           & BAUD-v1               & YES \\
\texttt{calibrator\_invents\_confidence}  & calibration policy           & MD-v1 + BAUD-v1       & YES \\
\texttt{decision\_proceeds\_anyway}       & decision policy              & BAUD-v1               & YES \\
\texttt{decision\_never\_proceeds}        & decision policy              & ERUR-v1               & YES \\
\texttt{actuation\_non\_safe\_reason}     & actuation policy             & BAUD-v1               & YES \\
\texttt{fpb\_threshold\_exceeded}         & verifier \texttt{max\_fire\_fraction} & FPB-v1     & YES \\
\bottomrule
\end{tabular}
\caption{Six bug categories, all detected by the verifier. Each
mini-smoke is engineered to break exactly one closed-loop
component; the verifier reports the expected property as VIOLATED
and the others as HOLDS. Reproducible via
\texttt{python -m project\_ghost.examples.violation\_matrix};
the script's exit code is 1 iff any false negative occurs.}
\label{tab:violation}
\end{table}

\subsection{Parametric policy evaluation}

To verify that the contracts are stable under variation of the
reference calibrator's parameters, we sweep three $(M,K)$ pairs
across three trace lengths. The result quantifies what an FMAS
reviewer expects to know before trusting a single all-HOLDS
verdict: the property set is not over-fit to the
$(M{=}4,K{=}2)$ point.

\begin{table}[h]
\centering\small
\begin{tabular}{l r r r r r c}
\toprule
\textbf{Policy} $(M,K)$ & $n$ & \textbf{MCAP (bytes)} & \textbf{Smoke (ms)} & \textbf{Verifier (ms)} & \textbf{Fire frac.} & \textbf{All HOLDS} \\
\midrule
(4, 2) reference   & 10  & 6\,552  & 15.3  & 20.8  & 0.60 & \checkmark \\
(4, 2) reference   & 50  & 18\,481 & 24.3  & 99.7  & 0.92 & \checkmark \\
(4, 2) reference   & 200 & 64\,699 & 100.5 & 405.8 & 0.98 & \checkmark \\
(3, 1) sensitive   & 10  & 6\,542  & 5.6   & 20.6  & 0.70 & \checkmark \\
(3, 1) sensitive   & 50  & 18\,475 & 34.0  & 100.2 & 0.94 & \checkmark \\
(3, 1) sensitive   & 200 & 64\,717 & 102.7 & 399.1 & 0.99 & \checkmark \\
(5, 3) lax         & 10  & 6\,553  & 5.6   & 20.6  & 0.50 & \checkmark \\
(5, 3) lax         & 50  & 18\,484 & 23.4  & 100.7 & 0.90 & \checkmark \\
(5, 3) lax         & 200 & 64\,700 & 104.4 & 406.2 & 0.98 & \checkmark \\
\bottomrule
\end{tabular}
\caption{Parametric policy sweep: three $(M,K)$ pairs $\times$
three trace lengths $n \in \{10, 50, 200\}$. All nine combinations
pass all five contracts. Verifier runtime is linear in trace
length; MCAP SHA-256 is byte-identical across replicate runs of
the same policy/$n$.}
\label{tab:parametric}
\end{table}

\subsection{Policy-comparison matrix (ERUR-v1 vs ERUR-v2)}

The verifier is policy-agnostic in the sense that the
\texttt{CalibrationAdjustmentPolicy} Protocol is the only
contract any calibrator must satisfy. ERUR, however, has a
predicate that depends on what the policy itself classifies as
"drift absent". Table~\ref{tab:policy-comparison} reports the
verdict for three structurally distinct calibrators against both
the v1 statement (fixed to the reference predicate) and the v2
lifting (delegating to each policy's own drift criterion); the
discrepancy is the operational evidence that the v2 lift is
meaningful and not nominal.

\begin{table}[h]
\centering\small
\begin{tabular}{l c c c c c c}
\toprule
\textbf{Policy} & \textbf{BAUD} & \textbf{ERUR-v1} & \textbf{ERUR-v2} & \textbf{MD} & \textbf{RLB} & \textbf{FPB} \\
\midrule
\texttt{Mahalanobis}$(M{=}4,K{=}2)$ ref.   & OK & OK             & OK             & OK & OK & OK \\
\texttt{EWMA}$(\alpha{=}0.5,\min{=}3,\tau{=}0.3)$ & OK & \textbf{VIOL} & \textbf{OK} & OK & OK & OK \\
\texttt{PerAxisHysteresis}$(\mathrm{up}{=}3.0)$  & OK & \textbf{VIOL} & \textbf{OK} & OK & OK & OK \\
\bottomrule
\end{tabular}
\caption{Verifier executed unchanged on three structurally
distinct calibrators. ERUR-v1 (fixed to the reference predicate)
reports VIOL on the alternatives because they do not behave like
the reference --- not because they are unsafe; ERUR-v2
(policy-parametric) captures the policy-agnostic contract that
the comparison matrix claims. Both verdicts are produced by
\texttt{verify\_erur} and \texttt{verify\_erur\_v2}
respectively (both ship in v0.2.4, ADR-0040). The matrix is
regenerated by \texttt{docs/paper/scripts/compare\_policies.py}
on every CI push and emitted to
\texttt{docs/paper/outputs/policy\_comparison.json}.}
\label{tab:policy-comparison}
\end{table}

\subsection{Realistic-shape scenarios}

To probe whether the contracts generalise beyond the
sustained-drift / single-recovery shape of the basic smoke, we
add three scenarios whose ground-truth functions are modelled
after failure modes documented in the VIO/SLAM evaluation
literature. They are still synthetic but they are no longer
trivially shaped.

\begin{table}[h]
\centering\small
\begin{tabular}{l r c c c c c r}
\toprule
\textbf{Scenario} & \textbf{Cycles} & \textbf{BAUD} & \textbf{ERUR} & \textbf{MD} & \textbf{RLB} & \textbf{FPB} & \textbf{fire frac.} \\
\midrule
\texttt{gps\_denial}        & 50 & OK & OK & OK & OK & OK & 0.64 \\
\texttt{slow\_biased\_drift}& 50 & OK & OK & OK & OK & OK & 0.92 \\
\texttt{cascading\_failure} & 30 & OK & OK & OK & OK & OK & 0.60 \\
\bottomrule
\end{tabular}
\caption{Three drift profiles shaped after VIO/SLAM evaluation
literature failure modes: short sustained drift +recovery
(\texttt{gps\_denial}), low-magnitude chronic drift
(\texttt{slow\_biased\_drift}), three-phase multi-axis cascade
(\texttt{cascading\_failure}). All five contracts hold under the
reference calibration policy.}
\label{tab:realistic}
\end{table}

\subsection{RTAMT capability matrix}

RTAMT and Ghost answer different questions; this matrix is not a
race. It enumerates the capabilities each tool offers over the
same MCAP so that a reader who is familiar with one can locate
the other on the same axes without re-deriving the comparison.

\begin{table}[h]
\centering\small
\begin{tabularx}{\linewidth}{X c c}
\toprule
\textbf{Capability} & \textbf{Ghost v0.2.3} & \textbf{RTAMT 0.3.5} \\
\midrule
Native property language & Python predicate on Ghost MCAP schema & Signal temporal logic (STL) \\
Reads MCAP directly & Yes & No (caller extracts signals) \\
Counts of $K$-in-$W$ as a single formula & Yes (intrinsic) & No (auxiliary counter signals) \\
Robustness semantics & No (verdict-only) & Yes \\
Arbitrary user-defined STL & Out of scope & Yes \\
Bug-detection over Ghost's reference pipeline & Demonstrated systematically (Table~\ref{tab:violation}) & Requires user re-encoding as STL \\
Distribution & PyPI + OIDC-signed wheel & PyPI source \\
Per-property typed report (\texttt{holds}, \texttt{mcap\_sha256}, violation metadata) & Yes & No (robustness value per timestamp) \\
\bottomrule
\end{tabularx}
\caption{Capability comparison, evaluated against published
feature sets. The two tools are complementary: RTAMT is the right
choice for declarative STL over user-defined signals with
quantitative robustness; Ghost is the right choice for a
content-addressed, schema-aware CLI verifier over epistemic
contracts. Welcome correction from RTAMT maintainers on any row
a recent release may have closed.}
\label{tab:rtamt-bench}
\end{table}

\end{document}
